---
title: "Independent Project Markdown"
author: "Maggie and Kendal"
date: "2026-07-20"
output: html_document
---

# Project Overview: 

The goal of this project is two fold; first, to address possible sampling gaps within the methods of the TEMPEST IV sampling campaign and secondly, the quantify short-term impacts of saltwater exposure on TEMPEST focal tree species photosynthetic rate and associated leaf physiology.

In order to sample canopy data, branches need to be removed from the tree, re-cut in a bucket of saltwater, and brought to the IRGAs in the lab. There is an unavoidable time delay between when the sample is removed, when the sample has access to water, and when the sample begins measurement. We want to know whether any of these actions impacts the ability of the leaves on the sample branch to photosynthesize so that we can better understand the veg campaign data.  
  
Additionally, it is difficult to know (and therefore sample) exactly when the trees in the saltwater plot are actively uptaking saltwater, and what the direct effects of this uptake would be. However, by snipping a sample in a bucket of saltwater, we may be able to mimic this uptake.    

This project investigates whether 1. the **action of snipping** a branch off the main stem 2. the **time delay** for that branch accessing water, and 3. the **type of water** the branch is in (i.e. freshwater versus saltwater) impact the photosynthesis of the three focal TEMPEST species (tulip poplar, red maple, and beech). 

  
Data collection is divided into two cycles:

**Cycle A**  | **Cycle B**
------------- | -------------
Control *(attached to tree)* |Control *(attached to tree)*
Bucket *(snipped in a bucket of freshwater, no time delay)* |Fresh *(snipped in a bucket of freshwater, ~1 min delay)*
Detached *(snipped without providing any water source)* |Salt *(snipped in a bucket of saltwater, ~1 min delay)*

Each cycle runs for 75 minutes, logging data every 120 seconds. Cycles where a treatment is applied log one reading before the treatment.  


# Code Output: 
**This code:**  
 * inputs data from the instruments (as a .csv file)  
 * divides the data into the respective cycles based on log number  
 * adds a "cycle" identifier  
 * adds a "species" identifier  
 * adds a "logNum" identifier to start all cycles at time = 0, regardless of whether or not it was the first log of the day 
 * graphs data:   
    + as a standalone reading  
    + by cycle  
    + [overall by species?]  
    + [overall by cycle?]  


```{r setup, include = FALSE} 
knitr::opts_chunk$set(echo = TRUE) #this is from original markdown doc
```

```{r loading packages, include = FALSE}
library(lubridate)
library(ggplot2)
library(dplyr)
```

<<<<<<< Updated upstream:testing markdown.Rmd
<<<<<<< Updated upstream
```{r import data, include = FALSE}
UTDay1 <- read.csv("2026-07-21_UT.csv") #LITU1 detached, LITU1 salt, FAGR1 detached
NoNameDay1 <- read.csv("2026-07-21_NoName.csv") #LITU1 bucket, LITU1 fresh, FAGR1 bucket
PNNLDay1 <- read.csv("2026-07-21 PNNL Day 1.csv") #LITU1 control A, LITU1 control B, FAGR1 control A

UTDay2 <- read.csv("2026-07-22 UT Day 2.csv") #LITU2 detached, LITU2 salt, FAGR1 salt, blank = obs 
NoNameDay2 <- read.csv("2026-07-22 NoName Day 2.csv") #LITU2 bucket, LITU2 fresh, FAGR1 fresh, blank = obs 115
PNNLDay2 <- read.csv("2026-07-22 PNNL Day 2.csv") #LITU2 controlA, LITU2 controlB, FAGR2 controlB, blank = obs 113

UTDay3 <- read.csv("2026-07-24 UT Day 3.csv") #FAGR2 detached, salt, blank = 77
NoNameDay3 <- read.csv("2026-07-24 NoName (misnamed PNNL)Day 3.csv") #FAGR2 control A, control b, blank = NA
PNNLDay3 <- read.csv("2026-07-24 PNNL (misnamed NoName) Day 3.csv") #FAGR2 bucket, fresh, blank = 77

UTDay4 <- read.csv("2026-07-29 UT Day 4.csv") #FAGR3 detached, FAGR3 salt, LITU3 salt
NoNameDay4 <- read.csv("2026-07-29 NoName Day 4.csv") #FAGR3 bucket, fresh, LITU3 fresh
PNNLDay4 <- read.csv("2026-07-29 PNNL Day 4.csv") #FAGR3 controlA,B, LITU3 controlB

NoNameDay5 <- read.csv("2026-07-30 noname day 5.csv") #LITU3 controlA
UTDay5 <- read.csv("2026-07-30 ut day 5.csv") #LITU3 bucket
PNNLDay5 <- read.csv("2026-07-30 pnnl day 5.csv") #LITU3 detached
```

```{r seperate into cycles}
#could i make this a function?? does r do functions???
#I MADE IT A FUNCTION!!! (it works!!) YAY!!
cleanData <- function(dataSheet, cycleNumber, cycleType, spp, obsStart, obsEnd) {
  #dataSheet = data, cylceNumber = String, cycleType = String, spp = String, obsStart = int, obsEnd = int
  dataSheet %>% #grabs from correct data sheet
    filter(between(obs, obsStart, obsEnd)) %>% #filters to correct log numbers
    mutate(logNum = obs - (obsStart- 1)) %>% #makes it so logNum always starts on 1
    mutate(name = cycleNumber) %>% #(like LITU_A1 = tulip cycle a rep 1), to ID cycle #
    mutate(cycle = cycleType) %>% #adds cycle type
    mutate(species = spp) %>% #adds species
    select(logNum, A, gsw, date, Ci, cycle, species) #get data!
}
#metadata sheet?

#LITU
LITU_A1_control <- cleanData(PNNLDay1, "LITU_A1", "controlA", "LITU", 2, 39)
LITU_A1_detached <- cleanData(UTDay1, "LITU_A1", "detached", "LITU", 1, 38)
LITU_A1_bucket <- cleanData(NoNameDay1, "LITU_A1", "bucket", "LITU", 1, 38)

LITU_B1_control <- cleanData(PNNLDay1, "LITU_B1", "controlB", "LITU", 41, 77)
LITU_B1_fresh <- cleanData(NoNameDay1, "LITU_B1", "fresh", "LITU", 40, 77)
LITU_B1_salt <- cleanData(UTDay1, "LITU_B1", "salt", "LITU", 40, 77)

LITU_A2_control <- cleanData(PNNLDay2, "LITU_A2", "controlA", "LITU", 1, 38)
LITU_A2_detached <- cleanData(UTDay2, "LITU_A2", "detached", "LITU", 1, 38)
LITU_A2_bucket <- cleanData(NoNameDay2, "LITU_A2", "bucket", "LITU", 1, 38)

LITU_B2_control <- cleanData(PNNLDay2, "LITU_B2", "controlB", "LITU", 39, 75)
LITU_B2_fresh <- cleanData(NoNameDay2, "LITU_B2", "fresh", "LITU", 39, 76)
LITU_B2_salt <- cleanData(UTDay2, "LITU_B2", "salt", "LITU", 39, 76)

#need to add in both LITU remaining cycles

LITU_A3_control <- cleanData(NoNameDay5, "LITU_A3", "controlA", "LITU", 1, 37)
LITU_A3_detached <- cleanData(PNNLDay5, "LITU_A3", "detached", "LITU", 1, 38)
LITU_A3_bucket <- cleanData(UTDay5, "LITU_A3", "bucket", "LITU", 1, 38)

LITU_B3_control <- cleanData(PNNLDay4, "LITU_B3", "controlB", "LITU", 75, 111)
LITU_B3_fresh <- cleanData(NoNameDay4, "LITU_B3", "fresh", "LITU", 77, 114)
LITU_B3_salt <- cleanData(UTDay4, "LITU_B3", "salt", "LITU", 77, 114)


#FAGR
FAGR_A1_control <- cleanData(PNNLDay1, "FAGR_A1", "controlA", "FAGR", 79, 116)
FAGR_A1_detached <- cleanData(UTDay1, "FAGR_A1", "detached", "FAGR", 79, 116)
FAGR_A1_bucket <- cleanData(NoNameDay1, "FAGR_A1", "bucket", "FAGR", 80, 117)

FAGR_B1_control <- cleanData(PNNLDay2, "FAGR_B1", "controlB", "FAGR", 76, 112)
FAGR_B1_fresh <- cleanData(NoNameDay2, "FAGR_B1", "fresh", "FAGR", 77, 114)
FAGR_B1_salt <- cleanData(UTDay2, "FAGR_B1", "salt", "FAGR", 77, 114)

FAGR_A2_control <- cleanData(NoNameDay3, "FAGR_A2", "controlA", "FAGR", 1, 37)
FAGR_A2_detached <- cleanData(UTDay3, "FAGR_A2", "detached", "FAGR", 1, 38)
FAGR_A2_bucket <- cleanData(PNNLDay3, "FAGR_A2", "bucket", "FAGR", 1, 38)

FAGR_B2_control <- cleanData(NoNameDay3, "FAGR_B2", "controlB", "FAGR", 38, 74)
FAGR_B2_fresh <- cleanData(PNNLDay3, "FAGR_B2", "fresh", "FAGR", 39, 76)
FAGR_B2_salt <- cleanData(UTDay3, "FAGR_B2", "salt", "FAGR", 39, 76)

FAGR_A3_control <- cleanData(NoNameDay4, "FAGR_A3", "controlA", "FAGR", 1, 37)
FAGR_A3_detached <- cleanData(UTDay4, "FAGR_A3", "detached", "FAGR", 1, 38)
FAGR_A3_bucket <- cleanData(PNNLDay4, "FAGR_A3", "bucket", "FAGR", 1, 38)

FAGR_B3_control <- cleanData(NoNameDay4, "FAGR_B3", "controlB", "FAGR", 38, 74)
FAGR_B3_fresh <- cleanData(PNNLDay4, "FAGR_B3", "fresh", "FAGR", 39, 76)
FAGR_B3_salt <- cleanData(UTDay4, "FAGR_B3", "salt", "FAGR", 39, 76)
=======
```{r post-kendal meeting work, include = FALSE, message = FALSE}

#KAM tries to read in data using metadata sheet
=======
```{r read-in data, include = FALSE}

#Metadata-based read-in
>>>>>>> Stashed changes:ToD Markdown.Rmd
data_dir <- "Data"
metadata <- read.csv("TTD_metadata.csv", stringsAsFactors = FALSE)

#need to look into what this one is doing specifically
##guess: creating a list of csv files based on if it has ".csv" in the "data" files
csv_list <- list.files(
  path = data_dir,
  pattern = "\\.csv$",
  full.names = TRUE
)

<<<<<<< Updated upstream:testing markdown.Rmd
#this is my function just cleaned!
=======
>>>>>>> Stashed changes:ToD Markdown.Rmd
cleanData <- function(dataSheet, IRGAName, cycleAorB, cycleType, spp, obsStart, obsEnd, file_name, replicate) {
  dataSheet %>%
    filter(between(obs, obsStart, obsEnd)) %>%
    mutate(
      logNum   = obs - (obsStart - 1),
      cycle    = cycleAorB,
      treatment = cycleType,
      species  = spp,
      File     = file_name,
      IRGAName = IRGAName,
      replicate = replicate
    ) %>%
<<<<<<< Updated upstream:testing markdown.Rmd
    select(date, IRGAName, species, cycle, treatment, replicate, logNum, A, gsw, Ci)
  #add additional columns above if more data is desired from licor output
}

#taking all data from metadata/csvs, cleaning data, returning as one data frame
##oh!!! R does use return(), but it's not required!
read_and_clean_all <- function(metadata, data_dir = "Data") {
  
  metadata <- metadata %>%
    mutate(File = as.character(File))
  
  # For each row in metadata, read, clean, and return a tibble
  all_clean <- metadata %>%
    pmap_dfr(function(Date, Instrument, Species, Treatment, Cycle, Replicate, Start, End, File, Notes) {
      
      # Build full file path
      fullpath <- file.path(data_dir, File)
      
      # Read raw CSV
      raw <- readr::read_csv(fullpath, show_col_types = FALSE)
      
      # Apply cleaning based on metadata fields
      cleanData(
        dataSheet  = raw,
        IRGAName   = Instrument,  # use the metadata "Instrument" (PNNL/NoName/UT)
        cycleAorB  = Cycle,       # use the metadata "Cycle" (A/B/etc.)
        cycleType  = Treatment,   # use the metadata "Treatment"
        spp        = Species,     # use the metadata "Species"
        obsStart   = Start,
        obsEnd     = End,
        file_name  = File,
        replicate  = Replicate    #use the metadata "Replicate"
      )
      
    })
  
  return(all_clean)
}

cleaned_all <- read_and_clean_all(metadata, data_dir = "Data")

glimpse(cleaned_all)

#adjusted with blanks test?
#needs to: 
  #average UT, NoName, and PNNL's blanks to get adj. factor for A and gsw
  #subtract that value from each A or gsw value in respective rows
#my Qs: how to get it to know which columns to subtract from?

cleanedAdjustedAll = cleaned_all |> 
  #blankA = mean(#IRGA A col where cycle = Blank)
  #blankGSW = mean(#IRGA gsw col where cycle = Blank)
  #group_by(IRGAName)
    mutate(adjA = A - blankA) #subtract blankA from all A rows
    mutate(adjGSW = gsw - blankGSW)#subtract blankGSW from all gsw rows
  #}


#the one that would tell when/if it reaches 20% decrease
placeholderName = function(percent, ){
  cleaned_all |> 
    group_by(species, treatment, cycle, replicate) %>%
    arrange(logNum) |> 
      #for the cycle, iterate one index at a time.
      #if there are three consecutive "fail" at a time, return first index of fail
    
  #return logNum where first true if true, else return NA
}


relative <- cleaned_all %>%
  group_by(species, treatment, cycle) %>%
  arrange(logNum) %>%
  mutate(A0 = first(A),
         gsw0 = first(gsw),
         A_rel = A/A0,
         gsw_rel = gsw/gsw0)
#%>% to calculate means of relative assmilation and stomatal conductance
# add to this summarize statement (can also do sd, counts, etc)
# summarize(mean_A = mean(A_rel))

```

```{r import data OLD, include = FALSE}
# UTDay1 <- read.csv("2026-07-21_UT.csv") #LITU1 detached, LITU1 salt, FAGR1 detached
# NoNameDay1 <- read.csv("2026-07-21_NoName.csv") #LITU1 bucket, LITU1 fresh, FAGR1 bucket
# PNNLDay1 <- read.csv("2026-07-21 PNNL Day 1.csv") #LITU1 control A, LITU1 control B, FAGR1 control A
# 
# UTDay2 <- read.csv("2026-07-22 UT Day 2.csv") #LITU2 detached, LITU2 salt, FAGR1 salt, blank = obs 
# NoNameDay2 <- read.csv("2026-07-22 NoName Day 2.csv") #LITU2 bucket, LITU2 fresh, FAGR1 fresh, blank = obs 115
# PNNLDay2 <- read.csv("2026-07-22 PNNL Day 2.csv") #LITU2 controlA, LITU2 controlB, FAGR2 controlB, blank = obs 113
# 
# UTDay3 <- read.csv("2026-07-24 UT Day 3.csv") #FAGR2 detached, salt, blank = 77
# NoNameDay3 <- read.csv("2026-07-24 NoName (misnamed PNNL)Day 3.csv") #FAGR2 control A, control b, blank = NA
# PNNLDay3 <- read.csv("2026-07-24 PNNL (misnamed NoName) Day 3.csv") #FAGR2 bucket, fresh, blank = 77
# 
# UTDay4 <- read.csv("2026-07-29 UT Day 4.csv") #FAGR3 detached, FAGR3 salt, LITU3 salt
# NoNameDay4 <- read.csv("2026-07-29 NoName Day 4.csv") #FAGR3 bucket, fresh, LITU3 fresh
# PNNLDay4 <- read.csv("2026-07-29 PNNL Day 4.csv") #FAGR3 controlA,B, LITU3 controlB
# 
# NoNameDay5 <- read.csv("2026-07-30 noname day 5.csv") #LITU3 controlA
# UTDay5 <- read.csv("2026-07-30 ut day 5.csv") #LITU3 bucket
# PNNLDay5 <- read.csv("2026-07-30 pnnl day 5.csv") #LITU3 detached
# 
# UTDay8 <- read.csv("data/2026-08-11 UT Day 8.csv") #controls for ACRU2B/ACRU3A
# NoNameDay8 <- read.csv("data/2026-08-11 NoName Day 8.csv") #fresh and bucket
```

```{r seperate into cycles OLD, include = FALSE}
# #could i make this a function?? does r do functions???
# #I MADE IT A FUNCTION!!! (it works!!) YAY!!
# cleanData <- function(dataSheet, cycleNumber, cycleType, spp, obsStart, obsEnd) {
#   #dataSheet = data, cylceNumber = String, cycleType = String, spp = String, obsStart = int, obsEnd = int
#   dataSheet %>% #grabs from correct data sheet
#     filter(between(obs, obsStart, obsEnd)) %>% #filters to correct log numbers
#     mutate(logNum = obs - (obsStart- 1)) %>% #makes it so logNum always starts on 1
#     mutate(name = cycleNumber) %>% #(like LITU_A1 = tulip cycle a rep 1), to ID cycle #
#     mutate(cycle = cycleType) %>% #adds cycle type
#     mutate(species = spp) %>% #adds species
#     select(logNum, A, gsw, date, Ci, cycle, species) #get data!
# }
# #metadata sheet?
# 
# #LITU
# LITU_A1_control <- cleanData(PNNLDay1, "LITU_A1", "controlA", "LITU", 2, 39)
# LITU_A1_detached <- cleanData(UTDay1, "LITU_A1", "detached", "LITU", 1, 38)
# LITU_A1_bucket <- cleanData(NoNameDay1, "LITU_A1", "bucket", "LITU", 1, 38)
# 
# LITU_B1_control <- cleanData(PNNLDay1, "LITU_B1", "controlB", "LITU", 41, 77)
# LITU_B1_fresh <- cleanData(NoNameDay1, "LITU_B1", "fresh", "LITU", 40, 77)
# LITU_B1_salt <- cleanData(UTDay1, "LITU_B1", "salt", "LITU", 40, 77)
# 
# LITU_A2_control <- cleanData(PNNLDay2, "LITU_A2", "controlA", "LITU", 1, 38)
# LITU_A2_detached <- cleanData(UTDay2, "LITU_A2", "detached", "LITU", 1, 38)
# LITU_A2_bucket <- cleanData(NoNameDay2, "LITU_A2", "bucket", "LITU", 1, 38)
# 
# LITU_B2_control <- cleanData(PNNLDay2, "LITU_B2", "controlB", "LITU", 39, 75)
# LITU_B2_fresh <- cleanData(NoNameDay2, "LITU_B2", "fresh", "LITU", 39, 76)
# LITU_B2_salt <- cleanData(UTDay2, "LITU_B2", "salt", "LITU", 39, 76)
# 
# #need to add in both LITU remaining cycles
# 
# LITU_A3_control <- cleanData(NoNameDay5, "LITU_A3", "controlA", "LITU", 1, 37)
# LITU_A3_detached <- cleanData(PNNLDay5, "LITU_A3", "detached", "LITU", 1, 38)
# LITU_A3_bucket <- cleanData(UTDay5, "LITU_A3", "bucket", "LITU", 1, 38)
# 
# LITU_B3_control <- cleanData(PNNLDay4, "LITU_B3", "controlB", "LITU", 75, 111)
# LITU_B3_fresh <- cleanData(NoNameDay4, "LITU_B3", "fresh", "LITU", 77, 114)
# LITU_B3_salt <- cleanData(UTDay4, "LITU_B3", "salt", "LITU", 77, 114)
# 
# 
# #FAGR
# FAGR_A1_control <- cleanData(PNNLDay1, "FAGR_A1", "controlA", "FAGR", 79, 116)
# FAGR_A1_detached <- cleanData(UTDay1, "FAGR_A1", "detached", "FAGR", 79, 116)
# FAGR_A1_bucket <- cleanData(NoNameDay1, "FAGR_A1", "bucket", "FAGR", 80, 117)
# 
# FAGR_B1_control <- cleanData(PNNLDay2, "FAGR_B1", "controlB", "FAGR", 76, 112)
# FAGR_B1_fresh <- cleanData(NoNameDay2, "FAGR_B1", "fresh", "FAGR", 77, 114)
# FAGR_B1_salt <- cleanData(UTDay2, "FAGR_B1", "salt", "FAGR", 77, 114)
# 
# FAGR_A2_control <- cleanData(NoNameDay3, "FAGR_A2", "controlA", "FAGR", 1, 37)
# FAGR_A2_detached <- cleanData(UTDay3, "FAGR_A2", "detached", "FAGR", 1, 38)
# FAGR_A2_bucket <- cleanData(PNNLDay3, "FAGR_A2", "bucket", "FAGR", 1, 38)
# 
# FAGR_B2_control <- cleanData(NoNameDay3, "FAGR_B2", "controlB", "FAGR", 38, 74)
# FAGR_B2_fresh <- cleanData(PNNLDay3, "FAGR_B2", "fresh", "FAGR", 39, 76)
# FAGR_B2_salt <- cleanData(UTDay3, "FAGR_B2", "salt", "FAGR", 39, 76)
# 
# FAGR_A3_control <- cleanData(NoNameDay4, "FAGR_A3", "controlA", "FAGR", 1, 37)
# FAGR_A3_detached <- cleanData(UTDay4, "FAGR_A3", "detached", "FAGR", 1, 38)
# FAGR_A3_bucket <- cleanData(PNNLDay4, "FAGR_A3", "bucket", "FAGR", 1, 38)
# 
# FAGR_B3_control <- cleanData(NoNameDay4, "FAGR_B3", "controlB", "FAGR", 38, 74)
# FAGR_B3_fresh <- cleanData(PNNLDay4, "FAGR_B3", "fresh", "FAGR", 39, 76)
# FAGR_B3_salt <- cleanData(UTDay4, "FAGR_B3", "salt", "FAGR", 39, 76)
# 
# 
# 
# ACRU_B2_control <- cleanData(UTDay8, "ACRU_B2", "controlB", "ACRU", 1, 38)
# ACRU_A3_control <- cleanData(UTDay8, "ACRU_A3", "controlA", "ACRU", 39, 76)
# 
# ACRU_B2_fresh <- cleanData(NoNameDay8, "ACRU_B2", "fresh", "ACRU", 1, 38)
# ACRU_A3_bucket <- cleanData(NoNameDay8, "ACRU_B2", "bucket", "ACRU", 39, 76)
>>>>>>> Stashed changes
=======
    select(date, IRGAName, species, cycle, 
           treatment, replicate, logNum, A, gsw, Ci)
}

#taking all data from metadata/csvs, cleaning data, returning as one data frame
read_and_clean_all <- function(metadata, data_dir = "Data") {
  metadata <- metadata %>%
    mutate(File = as.character(File)) #taking file name, making String
  #For each row in metadata, read, clean, and return a tibble
  all_clean <- metadata %>%
    pmap_dfr(function(Date, Instrument, Species, Treatment, Cycle, Replicate, Start, End, File, Notes) {
      fullpath <- file.path(data_dir, File) # Build full file path
      raw <- readr::read_csv(fullpath, show_col_types = FALSE)  # Read raw CSV
      cleanData( #cleaning data to rename/only use needed cols
        dataSheet  = raw,
        IRGAName   = Instrument,  # use the metadata "Instrument" (PNNL/NoName/UT)
        cycleAorB  = Cycle,       # use the metadata "Cycle" (A/B, NA = blank)
        cycleType  = Treatment,   # use the metadata "Treatment" (Control, Fresh, Salt, etc)
        spp        = Species,     # use the metadata "Species" (ACRU, FAGR, LITU)
        obsStart   = Start,       # first log num
        obsEnd     = End,         # last log num
        file_name  = File,        # which .csv the data is from
        replicate  = Replicate    # use the metadata "Replicate"
      )
    })
  return(all_clean)
}

all_data <- read_and_clean_all(metadata, data_dir = "Data")
glimpse(all_data)
```

```{r clean data}

#quick look at blank data
all_data %>%
  filter(treatment == "Blank") %>%
  pivot_longer(cols = c("A", "gsw"), 
               names_to = "Type", 
               values_to = "Value") %>%
  ggplot() + geom_point(aes(logNum, Value)) + 
  facet_grid(Type ~ IRGAName)

#adjust based on blanks data
all_data |> 
  group_by(IRGAName, treatment) |> 
    filter(treatment == "Blank",
           logNum > 3) |>
  #this removes the major wiggles from the first few minutes
      summarize(offsetA = mean(A),
                errorA = sd(A),
             offsetGSW = mean(gsw),
             errorGSW = sd(gsw)) -> blanks
#blank correct the values
blanks |>
  ungroup() |> 
  select(IRGAName, offsetA, offsetGSW) |> 
  full_join(all_data, by = "IRGAName") |> 
  mutate(adjA = A - offsetA,
         adjGSW = gsw - offsetGSW) %>% #now we can drop the rows that have the blank data, so we don't have to remove NA's later on
  drop_na(cycle) -> corrected

ggplot(corrected, aes(adjA, stat = "identity")) + 
  geom_histogram() + facet_wrap(.~species)
ggplot(corrected, aes(adjGSW, stat = "identity")) + 
  geom_histogram() + facet_wrap(.~species)

#based on these graphs A of 10 and gsw of 0.2 seem like reasonable cutoffs for all species

#cleaning data step first (so we only have to do it once)
data <- corrected %>%
  filter(adjA <= 10,
         adjGSW <= 0.2) #remove outliers
  
#calculate relative values based on initial data
initials <- data %>%
  group_by(species, treatment, cycle, replicate) %>%
  arrange(logNum) %>%
  slice(1:5) %>%
  summarize(Ai = mean(adjA),
            GSWi = mean(adjGSW))

relative <- data %>%
  left_join(initials, by = c("species", "treatment", "cycle", "replicate")) %>%
  mutate(A_rel = adjA/Ai,
         gsw_rel = adjGSW/GSWi)

#K note 8.16: this chunk is a good candidate for deleting
#we'll use the blank-corrected data or the relative blank corrected
#for everything below

# relativeAdjusted <- adjusted_all |> 
#   group_by(species, treatment, cycle) |> 
#   arrange(logNum) |> 
#   mutate(A0 = first(adjA),
#          gsw0 = first(adjA),
#          A_rel = adjA/A0,
#          gsw_rel = adjGSW/gsw0) |> 
#   ungroup() |> 
#   filter(!is.na(cycle), A < 12) |> 
#   group_by(species, logNum, treatment, cycle) |> 
#   summarize(A_mean = mean(A_rel),
#             A_sd = sd(A_rel),
#             gsw_mean = mean(gsw_rel),
#             gsw_sd = sd(gsw_rel))
>>>>>>> Stashed changes:ToD Markdown.Rmd

```

# Graphs
<<<<<<< Updated upstream:testing markdown.Rmd

ideas: 
can i have it so this page shows a dropdown menu OR a "enter what you want to see" ?

<<<<<<< Updated upstream
=======
## with K: 
```{r k graphs, include = FALSE}

###
### K chunk from August 12th meeting
###

#cleaning data step first (so we only have to do it once)
cleaned_all %>%
  filter(! is.na(cleaned_all$cycle),
         A < 12) -> data

#making means dataframe for plotting
data %>%
  group_by(species, logNum, treatment, cycle) %>%
  summarize(A_mean = mean(A),
            A_sd = sd(A),
            gsw_mean = mean(gsw),
            gsw_sd = sd(gsw)) -> clean_means

#making rolled timeseries dataframe for plotting
data %>%
  group_by(species, treatment, cycle, replicate) %>%
  arrange(logNum) %>%
  mutate(A_roll = zoo::rollmean(A, k = 3, align = "left", fill = NA),
         gsw_roll = zoo::rollmean(gsw, k = 3, align = "left", fill = NA)) -> rolled

#and lastly, making an average of the rolled values for plotting
#may decide later that any or all of these are not useful
rolled %>%
=======
ideas: 
can i color code things more properly?
how to make graphs more readable?

## with K: 
```{r all data graphs, include = FALSE}
###
### K chunk from August 12th meeting
### Updated 8.16
###

#(i like having the new var name out front to easily find it later!)
#oops - i just changed that in the chunk above, I'll keep it on the left from here on out

#making means data frame for plotting
clean_means <- data %>%
  group_by(species, logNum, treatment, cycle) %>%
  summarize(A_mean = mean(adjA),
            A_sd = sd(adjA),
            gsw_mean = mean(adjGSW),
            gsw_sd = sd(adjGSW)) #-> clean_means

#making rolled time series data frame for plotting
#rolled average = plots the average of this and the next two (n=3), smooths curve
rolled <- data %>%
  group_by(species, treatment, cycle, replicate) %>%
  arrange(logNum) %>%
  mutate(A_roll = zoo::rollmean(adjA, k = 3, align = "left", fill = NA),
         gsw_roll = zoo::rollmean(adjGSW, k = 3, align = "left", fill = NA)) # -> rolled

#and lastly, making an average of the rolled values for plotting
#may decide later that any or all of these are not useful
rolled_means <- rolled %>%
>>>>>>> Stashed changes:ToD Markdown.Rmd
  ungroup() %>%
  group_by(species, logNum, treatment, cycle) %>%
  summarize(A_mean = mean(A_roll),
            A_sd = sd(A_roll),
            gsw_mean = mean(gsw_roll),
<<<<<<< Updated upstream:testing markdown.Rmd
            gsw_sd = sd(gsw_roll)) -> rolled_means

#a half-decent plot with errorbars
ggplot() + geom_line(data = rolled_means, aes(logNum, A_mean, color = treatment),
                     linewidth = 1.25) + 
  geom_point(data = clean_means, aes(logNum, A_mean, color = treatment), alpha = 0.5) + 
  geom_errorbar(data = clean_means, aes(x = logNum, ymin = A_mean - A_sd, ymax = A_mean + A_sd,
                                       color = treatment), alpha = 0.3) +
  geom_hline(yintercept = 0, color = 'red') +
  facet_grid(species ~ cycle, scales = "free_y") + theme_bw()

#K skipped this
# cleaned_all %>%
#  filter(!is.na(cycle),
#         replicate = 1) %>%
# ggplot(aes(logNum, gsw, color = treatment)) +
#   geom_point() + geom_line() + facet_grid(species ~ cycle, scales = "free_y") +
#   ggtitle("Example Stomatal Conductance Panel")

=======
            gsw_sd = sd(gsw_roll)) #-> rolled_means

#summary of data for Stomatal Conductance
ggplot() + geom_line(data = rolled_means, 
                     aes(logNum, gsw_mean, color = treatment),
                     linewidth = 1.25) + 
  geom_point(data = clean_means, 
             aes(logNum, gsw_mean, color = treatment), 
             alpha = 0.5) + 
  geom_ribbon(data = clean_means, 
              aes(x = logNum, 
                  ymin = gsw_mean - gsw_sd, 
                  ymax = gsw_mean + gsw_sd,
                  color = treatment, fill = treatment), alpha = 0.1) +
  geom_hline(yintercept = 0, color = 'red') +
  facet_grid(species ~ cycle, scales = "free_y") + theme_bw() +
  ggtitle("Stomatal Conductance Data")

#summary of data for Assimilation
ggplot() + geom_line(data = rolled_means, 
                     aes(logNum, A_mean, color = treatment),
                     linewidth = 1.25) + 
  geom_point(data = clean_means, 
             aes(logNum, A_mean, color = treatment), 
             alpha = 0.5) + 
  geom_ribbon(data = clean_means, 
              aes(x = logNum, 
                  ymin = A_mean - A_sd, 
                  ymax = A_mean + A_sd,
                  color = treatment, fill = treatment), alpha = 0.1) +
  geom_hline(yintercept = 0, color = 'red') +
  facet_grid(species ~ cycle, scales = "free_y") + theme_bw() +
  ggtitle("Assimilation Data")

```

```{r relative values, include = FALSE}

#simple, easily adjustable plot of the data
ggplot(data[data$species == "ACRU",], 
       aes(logNum, A, color = as.factor(replicate))) + 
  geom_point() + facet_wrap(.~treatment, scales = "free", nrow = 3)

#ACRU assimilation is especially jumpy,
#would benefit most from smoothing


#first a little formatting of the data
fplusb <- data %>%
  ungroup() %>%
  mutate(across("treatment", str_replace, "Bucket", "Fresh"))

fplusb_rel <- relative %>%
  ungroup() %>%
  #filter(species == "ACRU" & treatment != "Bucket" & replicate != "1" | species %in% c("LITU", "FAGR")) %>%
  mutate(across("treatment", str_replace, "Bucket", "Fresh")) %>%
      #mutate(across("treatment", str_replace, "Fresh", "1 Minute Delay"),
    #     across("treatment", str_replace, "Bucket", "No Delay")) %>%
  group_by(species, treatment, replicate) |> 
  arrange(logNum) |>
  mutate(A_roll = rollmean(A_rel, k = 3, align = "left", fill = NA),
         gsw_roll = rollmean(gsw_rel, k = 3, align = "left", fill = NA))

fplusb_rel_means <- fplusb_rel |>
  ungroup() |> 
  group_by(species, logNum, treatment) |> 
  summarize(A_mean = mean(A_roll),
            A_sd = sd(A_roll),
            gsw_mean = mean(gsw_roll),
            gsw_sd = mean(gsw_roll))

#A
#very noisy data in Fresh ACRU
ggplot(data = fplusb_rel_means) +
  geom_line(aes(logNum, A_mean, color = treatment),
            linewidth = 1.1) + 
  geom_ribbon(aes(x = logNum,
                  ymin = A_mean - A_sd,
                  ymax = A_mean + A_sd,
                  color = treatment, fill = treatment),
              alpha = 0.15, color = NA) +
  geom_hline(yintercept = 1, color = 'red', linestyle = 2) +
  facet_grid(rows = "species", scales = "free_y") + theme_bw()  +
  labs(x = "Log Number", y = "relative A") 

#gsw
ggplot(data = fplusb_rel_means) +
  geom_line(aes(logNum, gsw_mean, color = treatment),
            linewidth = 1.1) + 
  geom_ribbon(aes(x = logNum,
                  ymin = gsw_mean - gsw_sd,
                  ymax = gsw_mean + gsw_sd,
                  color = treatment, fill = treatment),
              alpha = 0.15, color = NA) +
  geom_hline(yintercept = 1, color = 'red', linestyle = 2) +
  facet_grid(rows = "species", scales = "free_y") + theme_bw()  +
  labs(x = "Log Number", y = "relative GSW") 


```

```{r Salt vs Dettached}
clean_meansSaltDetached <- data %>%
  group_by(species, logNum, treatment) %>%
  filter(treatment != "Fresh", treatment != "Bucket") |> 
  summarize(A_mean = mean(adjA),
            A_sd = sd(adjA),
            gsw_mean = mean(adjGSW),
            gsw_sd = sd(adjGSW)) |> 
  ungroup() 


saltDetachedOnly <- data |> 
  ungroup() |>
  group_by(treatment, species, logNum) |> 
  filter(treatment != "Fresh", treatment != "Bucket") |> 
  ungroup() |> 
  group_by(species, treatment) |> 
  mutate(A_roll = rollmean(adjA, k = 3, align = "left", fill = NA),
         gsw_roll = rollmean(adjGSW, k = 3, align = "left", fill = NA)) |> 
  ungroup() |> 
  group_by(species, logNum, treatment) |> 
  summarize(A_mean = mean(A_roll),
            A_sd = sd(A_roll),
            gsw_mean = mean(gsw_roll),
            gsw_sd = mean(gsw_roll))


ggplot() + geom_line(data = saltDetachedOnly, aes(logNum, A_mean, color = treatment),
                     linewidth = 1.25) + 
  geom_point(data = clean_meansSaltDetached, aes(logNum, A_mean, color = treatment), alpha = 0.5) + 
  geom_ribbon(data = clean_meansSaltDetached, aes(x = logNum, ymin = A_mean - A_sd, ymax = A_mean + A_sd,
                                       color = treatment, fill = treatment), alpha = 0.15, color = NA) +
  geom_hline(yintercept = 0, color = 'red') +
  facet_grid(rows = "species", scales = "free_y") + theme_bw()  +
  labs(x = "Log Number", y = "A") 

#gsw
ggplot() + geom_line(data = saltDetachedOnly, aes(logNum, gsw_mean, color = treatment),
                     linewidth = 1.25) + 
  geom_point(data = clean_meansSaltDetached, aes(logNum, gsw_mean, color = treatment), alpha = 0.5) + 
  geom_ribbon(data = clean_meansSaltDetached, aes(x = logNum, ymin = gsw_mean - gsw_sd, ymax = gsw_mean + gsw_sd,
                                       color = treatment, fill = treatment), alpha = 0.15, color = NA) +
  geom_hline(yintercept = 0, color = 'red') +
  facet_grid(rows = "species", scales = "free_y") + theme_bw()  +
  labs(x = "Log Number", y = "gsw") 
```


#Statistics
```{r Statistics, echo = FALSE}
>>>>>>> Stashed changes:ToD Markdown.Rmd

#a simple test of bucket effect

#isolate data for testing effect
<<<<<<< Updated upstream:testing markdown.Rmd
 data %>%
  ungroup() %>%
  group_by(treatment, species, replicate) %>%
  filter(cycle == "A",
         treatment != "Detached",
         logNum < 26 & logNum > 14) %>%
  summarize(A_mean = mean(A),
            gsw_mean = mean(gsw)) -> simple_bucket_test

#Effect on photosynthetic rate
simple_bucket1 <- aov(A_mean ~ treatment * species,
=======
simple_bucket_test <-  data %>%
  ungroup() %>%
  filter(treatment %in% c("Control", "Fresh")) %>%
  group_by(treatment, species, replicate) %>%
  filter(logNum < 26 & logNum > 14) %>%
  ungroup() %>%
  group_by(species, treatment) %>%
  summarize(A_mean = mean(adjA),
            gsw_mean = mean(adjGSW))

#Effect on photosynthetic rate
simple_bucket1 <- aov(A_mean ~ treatment + species,
>>>>>>> Stashed changes:ToD Markdown.Rmd
                    data = simple_bucket_test) 
summary(simple_bucket1)


#Effect on stomatal conductance
<<<<<<< Updated upstream:testing markdown.Rmd
simple_bucket2 <- aov(gsw_mean ~ treatment * species,
                    data = simple_bucket_test) 
summary(simple_bucket2)

```

>>>>>>> Stashed changes
## one leaf: 
```{r graph one sample, echo=FALSE} 
#Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
# i want to make it a function again.
plotOneLeaf <- function(data, title = "No title provided", lowerBound = 0, upperBound = 4){
  #data = specific one you want to plot
  ggplot(data, aes(x = logNum, y = A)) +
    geom_line() + 
    geom_point() + 
    theme_bw() + 
    ggtitle(title) + 
    coord_cartesian(ylim = c(lowerBound, upperBound))
}

plotControl1 <- plotOneLeaf(control1)
plotControl1

plotOneLeaf(control2)

#from original code
#ggplot(PNNLData, aes(x= logNum, y = A)) + geom_line() + geom_point() + ggtitle("Control over time (both cycles)")
```


## one cycle: 
```{r graph one cycle, echo=FALSE}

##Functions ----
plotOneCycleA <- function(data1, data2, data3, title = "No title provided", lowerBound = -1, upperBound = 10) {
  #data1,2,3 = dataset, title = String, bounds = num (default 0:4)
  ggplot() + 
    geom_line(data = data1, aes(x = logNum, y = A, color = cycle)) +
    geom_line(data = data2, aes(x = logNum, y = A, color = cycle)) +
    geom_line(data = data3, aes(x = logNum, y = A, color = cycle)) +
    geom_point(data = data1, aes(x = logNum, y = A, color = cycle)) +
    geom_point(data = data2, aes(x = logNum, y = A, color = cycle)) +
    geom_point(data = data3, aes(x = logNum, y = A, color = cycle)) +
    theme_bw() +
    ggtitle(title) + 
    coord_cartesian(ylim = c(lowerBound, upperBound))
} 

plotOneCycleGSW <- function(data1, data2, data3, title = "No title provided", lowerBound = 0, upperBound = 0.3) { 
  #NOT WOKRING- WHY?
  #data1,2,3 = dataset, title = String, bounds = num (default -1:1)
  ggplot() + 
    geom_line(data = data1, aes(x = logNum, y = gsw, color = cycle)) +
    geom_line(data = data2, aes(x = logNum, y = gsw, color = cycle)) +
    geom_line(data = data3, aes(x = logNum, y = gsw, color = cycle)) +
    geom_point(data = data1, aes(x = logNum, y = gsw, color = cycle)) +
    geom_point(data = data2, aes(x = logNum, y = gsw, color = cycle)) +
    geom_point(data = data3, aes(x = logNum, y = gsw, color = cycle)) +
    theme_bw() +
    ggtitle(title) + 
    coord_cartesian(ylim = c(lowerBound, upperBound))
} 
#Debug: should make it so that cycle color is consistent regardless of input order. how to?

#LITU----
LITU_A1_graphA <- plotOneCycleA(LITU_A1_control, LITU_A1_bucket, LITU_A1_detached, "LITU A1", 0, 8)
LITU_A1_graphA
LITU_A1_graphGSW <- plotOneCycleGSW(LITU_A1_bucket, LITU_A1_control, LITU_A1_detached, "LITU A1", 0, 0.2)
LITU_A1_graphGSW

LITU_B1_graphA <- plotOneCycleA(LITU_B1_control, LITU_B1_fresh, LITU_B1_salt, "LITU B1", 0, 8)
LITU_B1_graphA
LITU_B1_graphGSW <- plotOneCycleGSW(LITU_B1_control, LITU_B1_fresh, LITU_B1_salt, "LITU B1", 0, 0.3)
LITU_B1_graphGSW

LITU_A2_graphA <- plotOneCycleA(LITU_A2_control, LITU_A2_bucket, LITU_A2_detached, "LITU A2", -1, 8)
LITU_A2_graphA
LITU_A2_graphGSW <- plotOneCycleGSW(LITU_A2_control, LITU_A2_bucket, LITU_A2_detached, "LITU A2", 0, 0.3)
LITU_A2_graphGSW

LITU_B2_graphA <- plotOneCycleA(LITU_B2_control, LITU_B2_fresh, LITU_B2_salt, "LITU B2", -1, 8)
LITU_B2_graphA
LITU_B2_graphGSW <- plotOneCycleGSW(LITU_B2_control, LITU_B2_fresh, LITU_B2_salt, "LITU B2", 0, 0.3)
LITU_B2_graphGSW

LITU_A3_graphA <- plotOneCycleA(LITU_A3_control, LITU_A3_detached, LITU_A3_bucket, "LITU A3", -1, 8)
LITU_A3_graphA
LITU_A3_graphGSW <- plotOneCycleGSW(LITU_A3_control, LITU_A3_detached, LITU_A3_bucket, "LITU A3", 0, 0.3)
LITU_A3_graphGSW

LITU_B3_graphA <- plotOneCycleA(LITU_B3_control, LITU_B3_fresh, LITU_B3_salt, "LITU B3", -1, 8)
LITU_B3_graphA
LITU_B3_graphGSW <- plotOneCycleGSW(LITU_B3_control, LITU_B3_fresh, LITU_B3_salt, "LITU B3", 0, 0.1)
LITU_B3_graphGSW

#FAGR ----
FAGR_A1_graphA <- plotOneCycleA(FAGR_A1_bucket, FAGR_A1_control, FAGR_A1_detached, "FAGR A1", -1, 7)
FAGR_A1_graphA
FAGR_A1_graphGSW <- plotOneCycleGSW(FAGR_A1_bucket, FAGR_A1_control, FAGR_A1_detached, "FAGR A1", 0, 0.05)
FAGR_A1_graphGSW

FAGR_B1_graphA <- plotOneCycleA(FAGR_B1_control, FAGR_B1_fresh, FAGR_B1_salt, "FAGR B1", -1, 7)
FAGR_B1_graphA
FAGR_B1_graphGSW <- plotOneCycleGSW(FAGR_B1_control, FAGR_B1_fresh, FAGR_B1_salt, "FAGR B1", 0, 0.1)
FAGR_B1_graphGSW

FAGR_A2_graphA <- plotOneCycleA(FAGR_A2_bucket, FAGR_A2_control, FAGR_A2_detached, "FAGR A2", -1, 5)
FAGR_A2_graphA 
FAGR_A2_graphGSW <- plotOneCycleGSW(FAGR_A2_bucket, FAGR_A2_control, FAGR_A2_detached, "FAGR A2", 0, 0.1)
FAGR_A2_graphGSW

FAGR_B2_graphA <- plotOneCycleA(FAGR_B2_control, FAGR_B2_fresh, FAGR_B2_salt, "FAGR B2", -1, 5)
FAGR_B2_graphA
FAGR_B2_graphGSW <- plotOneCycleGSW(FAGR_B2_control, FAGR_B2_fresh, FAGR_B2_salt, "FAGR B2", 0, 0.1)
FAGR_B2_graphGSW

FAGR_A3_graphA <- plotOneCycleA(FAGR_A3_bucket, FAGR_A3_control, FAGR_A3_detached, "FAGR A3", 0, 6)
FAGR_A3_graphA 
FAGR_A3_graphGSW <- plotOneCycleGSW(FAGR_A3_bucket, FAGR_A3_control, FAGR_A3_detached, "FAGR A3", 0, 0.1)
FAGR_A3_graphGSW

FAGR_B3_graphA <- plotOneCycleA(FAGR_B3_control, FAGR_B3_fresh, FAGR_B3_salt, "FAGR B3", 0, 6)
FAGR_B3_graphA
FAGR_B3_graphGSW <- plotOneCycleGSW(FAGR_B3_control, FAGR_B3_fresh, FAGR_B3_salt, "FAGR B3", 0, 0.1)
FAGR_B3_graphGSW

#ACRU ----


#----
#from original code
#ggplot() + 
 # geom_line(data = control2, aes(x = logNum, y = A, color = cycle)) +
  #geom_line(data = fresh, aes(x = logNum, y = A, color = cycle)) +
 # geom_line(data = salt, aes(x = logNum, y = A, color = cycle))+ 
 # theme_bw() + 
 # ggtitle("Cycle 2 (buckets)") + 
 # coord_cartesian(ylim = c(0, 4))
```


## by species?
``` {r graph by species test, echo = FALSE, include = FALSE}

#need to combine species into one variable
  # c can smash them together, but doesn't really help?
#how to? 
  #new variable called [species][treatment]Combined[A/gsw]
  #maybe it takes [col] from [data frame], adds together and /3 for each value in the row?
  #no, needs to average! (some control's don't have 3 obs bc i got lazy)


#THIS IS VERY MESSY. CLEAN LATER


#new idea: make new var ([tree]_[treatment]) with obs column (1:38), A and gsw for each replicate, and the avg of the three

#i was having trouble with pipe logic so i just made this a separate function
mergeData <- function(data1, data2, data3) {
  halfMerge = merge(data1, data2, by = "logNum", all.x = TRUE) 
  merge(halfMerge, data3, by = "logNum", all.x = TRUE)
}

#THSI DOES NOT WORK
combineData <- function(data1, data2, data3) {
   data = mergeData(data1, data2, data3) %>%  
   rename(A1 = A.x) %>% 
   rename(A2 = A.y) %>% 
   rename(A3 = A) %>% 
   rename(gsw1 = gsw.x) %>% 
   rename(gsw2 = gsw.y) %>% 
   rename(gsw3 = gsw) %>% 
   select(logNum, A1, A2, A3, gsw1, gsw2, gsw3, species, cycle) 
   #get the averages here
   mutate(data, avgA = rowMeans(select(data, c(A1, A2, A3)), na.rm = TRUE)) %>%  
   mutate(data, avgGSW = rowMeans(select(data, c(gsw1, gsw2, gsw3)), na.rm = TRUE))
}

LITUcontrolA <- combineData(LITU_A1_control, LITU_A2_control, LITU_A3_control)
LITUbucket <- combineData(LITU_A1_bucket, LITU_A2_bucket, LITU_A3_bucket)
LITUdetached <- combineData(LITU_A1_detached, LITU_A2_detached, LITU_A3_detached)
LITUcontrolB <- combineData(LITU_B1_control, LITU_B2_control, LITU_B3_control)
LITUsalt <- combineData(LITU_B1_salt, LITU_B2_salt, LITU_B3_salt)
LITUfresh <- combineData(LITU_B1_fresh, LITU_B2_fresh, LITU_B3_fresh)

FAGRcontrolA <- combineData(FAGR_A1_control, FAGR_A2_control, FAGR_A3_control)
FAGRbucket <- combineData(FAGR_A1_bucket, FAGR_A2_bucket, FAGR_A3_bucket)
FAGRdetached <- combineData(FAGR_A1_detached, FAGR_A2_detached, FAGR_A3_detached)
FAGRcontrolB <- combineData(FAGR_B1_control, FAGR_B2_control, FAGR_B3_control)
FAGRsalt <- combineData(FAGR_B1_salt, FAGR_B2_salt, FAGR_B3_salt)
FAGRfresh <- combineData(FAGR_B1_fresh, FAGR_B2_fresh, FAGR_B3_fresh)


plotAvgA <- function(data1, data2, data3, title = "No title provided", lowerBound = -1, upperBound = 10) {
  #data1,2,3 = dataset, title = String, bounds = num (default 0:4)
  ggplot() + 
    geom_line(data = subset(data1, !is.na(avgA)), aes(x = logNum, y = avgA, color = cycle)) +
    geom_line(data = subset(data2, !is.na(avgA)), aes(x = logNum, y = avgA, color = cycle)) +
    geom_line(data = subset(data3, !is.na(avgA)), aes(x = logNum, y = avgA, color = cycle)) +
    geom_point(data = subset(data1, !is.na(avgA)), aes(x = logNum, y = avgA, color = cycle)) +
    geom_point(data = subset(data2, !is.na(avgA)), aes(x = logNum, y = avgA, color = cycle)) +
    geom_point(data = subset(data3, !is.na(avgA)), aes(x = logNum, y = avgA, color = cycle)) +
    theme_bw() +
    ggtitle(title) + 
    coord_cartesian(ylim = c(lowerBound, upperBound))
}

plotAvgGSW <- function(data1, data2, data3, title = "No title provided", lowerBound = 0, upperBound = 0.2) {
  #data1,2,3 = dataset, title = String, bounds = num (default 0:4)
  ggplot() + 
    geom_line(data = subset(data1, !is.na(avgA)), aes(x = logNum, y = avgGSW, color = cycle)) +
    geom_line(data = subset(data2, !is.na(avgA)), aes(x = logNum, y = avgGSW, color = cycle)) +
    geom_line(data = subset(data3, !is.na(avgA)), aes(x = logNum, y = avgGSW, color = cycle)) +
    geom_point(data = subset(data1, !is.na(avgA)), aes(x = logNum, y = avgGSW, color = cycle)) +
    geom_point(data = subset(data2, !is.na(avgA)), aes(x = logNum, y = avgGSW, color = cycle)) +
    geom_point(data = subset(data3, !is.na(avgA)), aes(x = logNum, y = avgGSW, color = cycle)) +
    theme_bw() +
    ggtitle(title) + 
    coord_cartesian(ylim = c(lowerBound, upperBound))
}


#clean this up later
LITU_A_avgA <- plotAvgA(LITUcontrolA, LITUbucket, LITUdetached)
LITU_A_avgA
LITU_B_avgA <- plotAvgA(LITUcontrolB, LITUfresh, LITUsalt)
LITU_B_avgA

LITU_A_avgGSW <- plotAvgGSW(LITUcontrolA, LITUbucket, LITUdetached, "LITU A gsw", 0, 0.2)
LITU_A_avgGSW
LITU_B_avgGSW <- plotAvgGSW(LITUcontrolB, LITUfresh, LITUsalt, "LITU B gsw", 0, 0.2)
LITU_B_avgGSW

ggplot() + 
  geom_line(data = LITUcontrolA, aes(x = logNum, y = avgA, color = cycle)) +
  geom_line(data = LITUcontrolB, aes(x = logNum, y = avgA, color = cycle)) + 
  geom_line(data = LITUbucket, aes(x = logNum, y = avgA, color = cycle)) +
  geom_line(data = LITUdetached, aes(x = logNum, y = avgA, color = cycle)) +
  geom_line(data = LITUfresh, aes(x = logNum, y = avgA, color = cycle)) + 
  geom_line(data = LITUsalt, aes(x = logNum, y = avgA, color = cycle)) +
  geom_point(data = LITUcontrolA, aes(x = logNum, y = avgA, color = cycle)) +
  geom_point(data = LITUcontrolB, aes(x = logNum, y = avgA, color = cycle)) + 
  geom_point(data = LITUbucket, aes(x = logNum, y = avgA, color = cycle)) +
  geom_point(data = LITUdetached, aes(x = logNum, y = avgA, color = cycle)) +
  geom_point(data = LITUfresh, aes(x = logNum, y = avgA, color = cycle)) + 
  geom_point(data = LITUsalt, aes(x = logNum, y = avgA, color = cycle)) +
=======
simple_bucket2 <- aov(gsw_mean ~ treatment + species,
                    data = simple_bucket_test) 
summary(simple_bucket2)

#hey m!
#bring any other anova style tests of the effect of cutting into this section!

```

```{r KM analysis}

# #the one that would tell when/if it reaches x% decrease ATTEMPT 1 ----
#K here, let's simplify, first let's make the check values and confirm that all looks good
getTimeDecrease <- fplusb |> 
  group_by(species, treatment, cycle, replicate) %>%
  arrange(logNum) %>%
  mutate(aCheck = first(adjA) * 0.8, #0.8 = 20% decrease
         gswCheck = first(adjGSW) * 0.8) %>%
  select(species, treatment, replicate, aCheck, gswCheck) %>%
  distinct()

#And now we can create a new column based on that value after we rejoin
km_data <- fplusb %>%
  left_join(getTimeDecrease, by = c("species", "treatment", "cycle", "replicate")) %>%
  mutate(Astatus = if_else(adjA > aCheck, "happy", "sad"),
        GSWstatus = if_else(adjGSW > gswCheck, "happy", "sad"),
        km_status1 = case_when(Astatus == "happy" ~ 0,
                               Astatus == "sad" ~ 1),
        km_status2 = case_when(GSWstatus == "happy" ~ 0,
                               GSWstatus == "sad" ~ 1))

library(survival)
library(survminer)
library(ranger)
library(ggfortify)
library(MuMIn)

km_data1 <- km_data %>%
  filter(treatment %in% c("Control", "Fresh"))

km_trt_fit <- survfit(Surv(logNum, km_status1) ~ treatment, data=km_data1)
summary(km_trt_fit)
p1 = autoplot(km_trt_fit) +
  labs(x='Number of Logs',
       y='Percent surviving (± 95%CI)',
       color='Handling',
       fill='Handling') +
>>>>>>> Stashed changes:ToD Markdown.Rmd
  theme_bw() +
  scale_color_manual(values=c('#2ca25f','#b30000')) +
  scale_fill_manual(values=c('#2ca25f','#b30000'))

```

<<<<<<< Updated upstream:testing markdown.Rmd
## by treatment?
=======
```{r Tempest in a Bucket}
km_data2 <- km_data %>%
  filter(treatment %in% c("Fresh", "Salt"))

km_trt_fit2 <- survfit(Surv(logNum, km_status1) ~ treatment, data=km_data2)
summary(km_trt_fit2)
p2 = autoplot(km_trt_fit2) +
  labs(x='Number of Logs',
       y='Percent surviving (± 95%CI)',
       color='Handling',
       fill='Handling') +
  theme_bw() +
  scale_color_manual(values=c('#2ca25f','#b30000')) +
  scale_fill_manual(values=c('#2ca25f','#b30000'))

```

<!-- ## by species -->
<!-- ``` {r graph by species NEW, echo = FALSE, include = fALSE}  -->

>>>>>>> Stashed changes:ToD Markdown.Rmd



<!-- ``` -->
