---
title: "Independent Project Markdown MESSY VERSION NOT CLEAN"
subtitle: "MAGGIE CLEAN THE CODE IN A DOCUMENT COPY"
author: "Maggie and Kendal"
date: "2026-07-20"
output: html_document
---

# Project Overview: 

The goal of this project is to address possible sampling gaps within the methods of the TEMPEST IV sampling campaign.

In order to sample canopy data, branches need to be removed from the tree, re-cut in a bucket of saltwater, and brought to the IRGAs in the lab. There is an unavoidable time delay between when the sample is removed, when the sample has access to water, and when the sample begins measurement. We want to know whether any of these actions impacts the ability of the leaves on the sample branch to photosynthesize so that we can better understand the veg campaign data.  
  
Additionally, it is difficult to know (and therefore sample) exactly when the trees in the saltwater plot are actively uptaking saltwater, and what the direct effects of this uptake would be. However, by snipping a sample in a bucket of saltwater, we may be able to mimic this uptake.    

This project investigates whether 1. the **action of snipping** a branch off the main stem 2. the **time delay** for that branch accessing water, and 3. the **type of water** the branch is in (i.e. freshwater versus saltwater) impact the photosynthesis of the three focal TEMPEST species (tulip poplar, red maple, and beech). 

  
Data collection is divided into two cycles:

**Cycle A**  | **Cycle B**
------------- | -------------
Control *(attached to tree)* |Control *(attached to tree)*
Bucket *(snipped in a bucket of freshwater, no time delay)* |Fresh *(snipped in a bucket of freshwater, ~1 min delay)*
Detached *(snipped without providing any water source)* |Salt *(snipped in a bucket of saltwater, ~1 min delay)*

Each cycle runs for 75 minutes, logging data every 120 seconds. Cycles where a treatment is applied log one reading before the treatment.  


# Code Output: 
**This code:**  
 * inputs data from the instruments (as a .csv file)  
 * divides the data into the respective cycles based on log number  
 * adds a "cycle" identifier  
 * adds a "species" identifier  
 * adds a "logNum" identifier to start all cycles at time = 0, regardless of whether or not it was the first log of the day 
 * graphs data:   
    + as a standalone reading  
    + by cycle  
    + [overall by species?]  
    + [overall by cycle?]  


```{r setup, include=FALSE} 
knitr::opts_chunk$set(echo = TRUE) #this is from original markdown doc
```

```{r loading packages, include = FALSE}
library(lubridate)
library(ggplot2)
library(dplyr)
```

```{r import data, include = FALSE}
UTTest <- read.csv("2026-07-16-0925_UTProcTest_cleaned.csv") #bucket, fresh
NoNameTest <- read.csv("2026-07-16-0824_NoNameProcTest_cleaned.csv") #detached, salt
PNNLTest <- read.csv("2026-07-16_PNNLProcTest_cleaned.csv") #control a, control b

UTDay1 <- read.csv("2026-07-21_UT.csv") #LITU1 detached, LITU1 salt, FAGR1 detached
NoNameDay1 <- read.csv("2026-07-21_NoName.csv") #LITU1 bucket, LITU1 fresh, FAGR1 bucket
PNNLDay1 <- read.csv("2026-07-21 PNNL Day 1.csv") #LITU1 control A, LITU1 control B, FAGR1 control A

UTDay2 <- read.csv("2026-07-22 UT Day 2.csv") #LITU2 detached, LITU2 salt, FAGR1 salt, blank = obs 
NoNameDay2 <- read.csv("2026-07-22 NoName Day 2.csv") #LITU2 bucket, LITU2 fresh, FAGR1 fresh, blank = obs 115
PNNLDay2 <- read.csv("2026-07-22 PNNL Day 2.csv") #LITU2 controlA, LITU2 controlB, FAGR2 controlB, blank = obs 113

UTDay3 <- read.csv("2026-07-24 UT Day 3.csv") #FAGR2 detached, salt, blank = 77
NoNameDay3 <- read.csv("2026-07-24 NoName (misnamed PNNL)Day 3.csv") #FAGR2 control A, control b, blank = NA
PNNLDay3 <- read.csv("2026-07-24 PNNL (misnamed NoName) Day 3.csv") #FAGR2 bucket, fresh, blank = 77

UTDay4 <- read.csv("2026-07-29 UT Day 4.csv") #FAGR3 detached, FAGR3 salt, LITU3 salt
NoNameDay4 <- read.csv("2026-07-29 NoName Day 4.csv") #FAGR3 bucket, fresh, LITU3 fresh
PNNLDay4 <- read.csv("2026-07-29 PNNL Day 4.csv") #FAGR3 controlA,B, LITU3 controlB
```

```{r seperate into cycles}

#could i make this a function?? does r do functions???
#I MADE IT A FUNCTION!!! (it works!!) YAY!!
cleanData <- function(dataSheet, cycleNumber, cycleType, spp, obsStart, obsEnd) {
  #dataSheet = data, cylceNumber = String, cycleType = String, spp = String, obsStart = int, obsEnd = int
  dataSheet %>% #grabs from correct data sheet
    filter(between(obs, obsStart, obsEnd)) %>% #filters to correct log numbers
    mutate(logNum = obs - (obsStart- 1)) %>% #makes it so logNum always starts on 1
    mutate(name = cycleNumber) %>% #(like LITU_A1 = tulip cycle a rep 1), to ID cycle #
    mutate(cycle = cycleType) %>% #adds cycle type
    mutate(species = spp) %>% #adds species
    select(logNum, A, gsw, date, Ci, cycle, species) #get data!
}


#metadata sheet?

#LITU
LITU_A1_control <- cleanData(PNNLDay1, "LITU_A1", "controlA", "LITU", 2, 39)
LITU_A1_detached <- cleanData(UTDay1, "LITU_A1", "detached", "LITU", 1, 38)
LITU_A1_bucket <- cleanData(NoNameDay1, "LITU_A1", "bucket", "LITU", 1, 38)

LITU_B1_control <- cleanData(PNNLDay1, "LITU_B1", "controlB", "LITU", 41, 77)
LITU_B1_fresh <- cleanData(NoNameDay1, "LITU_B1", "fresh", "LITU", 40, 77)
LITU_B1_salt <- cleanData(UTDay1, "LITU_B1", "salt", "LITU", 40, 77)

LITU_A2_control <- cleanData(PNNLDay2, "LITU_A2", "controlA", "LITU", 1, 38)
LITU_A2_detached <- cleanData(UTDay2, "LITU_A2", "detached", "LITU", 1, 38)
LITU_A2_bucket <- cleanData(NoNameDay2, "LITU_A2", "bucket", "LITU", 1, 38)

LITU_B2_control <- cleanData(PNNLDay2, "LITU_B2", "controlB", "LITU", 39, 75)
LITU_B2_fresh <- cleanData(NoNameDay2, "LITU_B2", "fresh", "LITU", 39, 76)
LITU_B2_salt <- cleanData(UTDay2, "LITU_B2", "salt", "LITU", 39, 76)




LIT


#FAGR
FAGR_A1_control <- cleanData(PNNLDay1, "FAGR_A1", "controlA", "FAGR", 79, 116)
FAGR_A1_detached <- cleanData(UTDay1, "FAGR_A1", "detached", "FAGR", 79, 116)
FAGR_A1_bucket <- cleanData(NoNameDay1, "FAGR_A1", "bucket", "FAGR", 80, 117)

FAGR_B1_control <- cleanData(PNNLDay2, "FAGR_B1", "controlB", "FAGR", 76, 112)
FAGR_B1_fresh <- cleanData(NoNameDay2, "FAGR_B1", "fresh", "FAGR", 77, 114)
FAGR_B1_salt <- cleanData(UTDay2, "FAGR_B1", "salt", "FAGR", 77, 114)

FAGR_A2_control <- cleanData(NoNameDay3, "FAGR_A2", "controlA", "FAGR", 1, 37)
FAGR_A2_detached <- cleanData(UTDay3, "FAGR_A2", "detached", "FAGR", 1, 38)
FAGR_A2_bucket <- cleanData(PNNLDay3, "FAGR_A2", "bucket", "FAGR", 1, 38)

FAGR_B2_control <- cleanData(NoNameDay3, "FAGR_B2", "controlB", "FAGR", 38, 74)
FAGR_B2_fresh <- cleanData(PNNLDay3, "FAGR_B2", "fresh", "FAGR", 39, 76)
FAGR_B2_salt <- cleanData(UTDay3, "FAGR_B2", "salt", "FAGR", 39, 76)

#check log numbers!!
FAGR_A3_control <- cleanData(NoNameDay4, "FAGR_A3", "controlA", "FAGR", 1, 37)
FAGR_A3_detached <- cleanData(UTDay4, "FAGR_A3", "detached", "FAGR", 1, 38)
FAGR_A3_bucket <- cleanData(PNNLDay4, "FAGR_A3", "bucket", "FAGR", 1, 38)

FAGR_B3_control <- cleanData(NoNameDay4, "FAGR_B3", "controlB", "FAGR", 38, 74)
FAGR_B3_fresh <- cleanData(PNNLDay4, "FAGR_B3", "fresh", "FAGR", 39, 76)
FAGR_B3_salt <- cleanData(UTDay4, "FAGR_B3", "salt", "FAGR", 39, 76)

```


# Graphs

ideas: 
can i have it so this page shows a dropdown menu OR a "enter what you want to see" ?

## one leaf: 
```{r graph one sample, echo=FALSE} 
#Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
# i want to make it a function again.
plotOneLeaf <- function(data, title = "No title provided", lowerBound = 0, upperBound = 4){
  #data = specific one you want to plot
  ggplot(data, aes(x = logNum, y = A)) +
    geom_line() + 
    geom_point() + 
    theme_bw() + 
    ggtitle(title) + 
    coord_cartesian(ylim = c(lowerBound, upperBound))
}

plotControl1 <- plotOneLeaf(control1)
plotControl1

plotOneLeaf(control2)

#from original code
#ggplot(PNNLData, aes(x= logNum, y = A)) + geom_line() + geom_point() + ggtitle("Control over time (both cycles)")
```


## one cycle: 
```{r graph one cycle, echo=FALSE}

plotOneCycleA <- function(data1, data2, data3, title = "No title provided", lowerBound = -1, upperBound = 10) {
  #data1,2,3 = dataset, title = String, bounds = num (default 0:4)
  ggplot() + 
    geom_line(data = data1, aes(x = logNum, y = A, color = cycle)) +
    geom_line(data = data2, aes(x = logNum, y = A, color = cycle)) +
    geom_line(data = data3, aes(x = logNum, y = A, color = cycle)) +
    geom_point(data = data1, aes(x = logNum, y = A, color = cycle)) +
    geom_point(data = data2, aes(x = logNum, y = A, color = cycle)) +
    geom_point(data = data3, aes(x = logNum, y = A, color = cycle)) +
    theme_bw() +
    ggtitle(title) + 
    coord_cartesian(ylim = c(lowerBound, upperBound))
} 

plotOneCycleGSW <- function(data1, data2, data3, title = "No title provided", lowerBound = 0, upperBound = 0.3) { 
  #NOT WOKRING- WHY?
  #data1,2,3 = dataset, title = String, bounds = num (default -1:1)
  ggplot() + 
    geom_line(data = data1, aes(x = logNum, y = gsw, color = cycle)) +
    geom_line(data = data2, aes(x = logNum, y = gsw, color = cycle)) +
    geom_line(data = data3, aes(x = logNum, y = gsw, color = cycle)) +
    geom_point(data = data1, aes(x = logNum, y = gsw, color = cycle)) +
    geom_point(data = data2, aes(x = logNum, y = gsw, color = cycle)) +
    geom_point(data = data3, aes(x = logNum, y = gsw, color = cycle)) +
    theme_bw() +
    ggtitle(title) + 
    coord_cartesian(ylim = c(lowerBound, upperBound))
} 
#Debug: should make it so that cycle color is consistent regardless of input order. how to?

plotOneCycleA(control1, bucket, detached, "Test A", 0, 4)
plotOneCycleA(control2, fresh, salt, "Test B", 0, 4)

LITU_A1_graphA <- plotOneCycleA(LITU_A1_control, LITU_A1_bucket, LITU_A1_detached, "LITU A1", 0, 8)
LITU_A1_graphA
LITU_A1_graphGSW <- plotOneCycleGSW(LITU_A1_bucket, LITU_A1_control, LITU_A1_detached, "LITU A1", 0, 0.2)
LITU_A1_graphGSW

LITU_B1_graphA <- plotOneCycleA(LITU_B1_control, LITU_B1_fresh, LITU_B1_salt, "LITU B1", 0, 8)
LITU_B1_graphA
LITU_B1_graphGSW <- plotOneCycleGSW(LITU_B1_control, LITU_B1_fresh, LITU_B1_salt, "LITU B1", 0, 0.3)
LITU_B1_graphGSW

LITU_A2_graphA <- plotOneCycleA(LITU_A2_control, LITU_A2_bucket, LITU_A2_detached, "LITU A2", -1, 8)
LITU_A2_graphA
LITU_A2_graphGSW <- plotOneCycleGSW(LITU_A2_control, LITU_A2_bucket, LITU_A2_detached, "LITU A2", 0, 0.3)
LITU_A2_graphGSW

LITU_B2_graphA <- plotOneCycleA(LITU_B2_control, LITU_B2_fresh, LITU_B2_salt, "LITU B2", -1, 8)
LITU_B2_graphA
LITU_B2_graphGSW <- plotOneCycleGSW(LITU_B2_control, LITU_B2_fresh, LITU_B2_salt, "LITU B2", 0, 0.3)
LITU_B2_graphGSW

FAGR_A1_graphA <- plotOneCycleA(FAGR_A1_bucket, FAGR_A1_control, FAGR_A1_detached, "FAGR A1", -1, 7)
FAGR_A1_graphA
FAGR_A1_graphGSW <- plotOneCycleGSW(FAGR_A1_bucket, FAGR_A1_control, FAGR_A1_detached, "FAGR A1", 0, 0.05)
FAGR_A1_graphGSW

FAGR_B1_graphA <- plotOneCycleA(FAGR_B1_control, FAGR_B1_fresh, FAGR_B1_salt, "FAGR B1", -1, 7)
FAGR_B1_graphA
FAGR_B1_graphGSW <- plotOneCycleGSW(FAGR_B1_control, FAGR_B1_fresh, FAGR_B1_salt, "FAGR B1", 0, 0.1)
FAGR_B1_graphGSW

FAGR_A2_graphA <- plotOneCycleA(FAGR_A2_bucket, FAGR_A2_control, FAGR_A2_detached, "FAGR A2", -1, 5)
FAGR_A2_graphA 
FAGR_A2_graphGSW <- plotOneCycleGSW(FAGR_A2_bucket, FAGR_A2_control, FAGR_A2_detached, "FAGR A2", 0, 0.1)
FAGR_A2_graphGSW

FAGR_B2_graphA <- plotOneCycleA(FAGR_B2_control, FAGR_B2_fresh, FAGR_B2_salt, "FAGR B2", -1, 5)
FAGR_B2_graphA
FAGR_B2_graphGSW <- plotOneCycleGSW(FAGR_B2_control, FAGR_B2_fresh, FAGR_B2_salt, "FAGR B2", 0, 0.1)
FAGR_B2_graphGSW



#from original code
#ggplot() + 
 # geom_line(data = control2, aes(x = logNum, y = A, color = cycle)) +
  #geom_line(data = fresh, aes(x = logNum, y = A, color = cycle)) +
 # geom_line(data = salt, aes(x = logNum, y = A, color = cycle))+ 
 # theme_bw() + 
 # ggtitle("Cycle 2 (buckets)") + 
 # coord_cartesian(ylim = c(0, 4))
```


## by species?
``` {r graph by species test, include = FALSE}

#need to combine species into one variable
  # c can smash them together, but doesn't really help?
#how to? 
  #new variable called [species][treatment]Combined[A/gsw]
  #maybe it takes [col] from [data frame], adds together and /3 for each value in the row?



speciesAverageATest <- function(data1, data2, data3) {
  (data1$A + data2$A + data3$A)/3
} #i think this should work? #yes works but will need to like. actually test when i get more data tonight

testAvg <-speciesAverageATest(LITU_B1_salt, LITU_B2_salt, LITU_B1_salt)
all.equal(testAvg,LITU_B1_salt$A)
testAvg

```

## by treatment?





