---
title: "Independent Project Markdown"
author: "Maggie and Kendal"
date: "2026-07-20"
output: html_document
---

# Project Overview: 

The goal of this project is to address possible sampling gaps within the methods of the TEMPEST IV sampling campaign.

In order to sample canopy data, branches need to be removed from the tree, re-cut in a bucket of saltwater, and brought to the IRGAs in the lab. There is an unavoidable time delay between when the sample is removed, when the sample has access to water, and when the sample begins measurement. We want to know whether any of these actions impacts the ability of the leaves on the sample branch to photosynthesize so that we can better understand the veg campaign data.  
  
Additionally, it is difficult to know (and therefore sample) exactly when the trees in the saltwater plot are actively uptaking saltwater, and what the direct effects of this uptake would be. However, by snipping a sample in a bucket of saltwater, we may be able to mimic this uptake.    

This project investigates whether 1. the **action of snipping** a branch off the main stem 2. the **time delay** for that branch accessing water, and 3. the **type of water** the branch is in (i.e. freshwater versus saltwater) impact the photosynthesis of the three focal TEMPEST species (tulip poplar, red maple, and beech). 

  
Data collection is divided into two cycles:

**Cycle A**  | **Cycle B**
------------- | -------------
Control *(attached to tree)* |Control *(attached to tree)*
Bucket *(snipped in a bucket of freshwater, no time delay)* |Fresh *(snipped in a bucket of freshwater, ~1 min delay)*
Detached *(snipped without providing any water source)* |Salt *(snipped in a bucket of saltwater, ~1 min delay)*

Each cycle runs for 75 minutes, logging data every 120 seconds. Cycles where a treatment is applied log one reading before the treatment.  
    
    
#STEPS: 
##make some graphs of each individual relative data (this beech, etc) and the summarized relative data (all beech, etc)

```{r setup, include=FALSE} 
knitr::opts_chunk$set(echo = TRUE) #this is from original markdown doc
```

```{r loading packages, include = FALSE}
library(lubridate)
library(ggplot2)
library(dplyr)
library(tidyverse)
library(zoo)
```

```{r Data read-in and cleaning, include = FALSE}

#Metadata-based read-in
data_dir <- "Data"
metadata <- read.csv("TTD_metadata.csv", stringsAsFactors = FALSE)

#need to look into what this one is doing specifically
##guess: creating a list of csv files based on if it has ".csv" in the "data" files
csv_list <- list.files(
  path = data_dir,
  pattern = "\\.csv$",
  full.names = TRUE
)

cleanData <- function(dataSheet, IRGAName, cycleAorB, cycleType, spp, obsStart, obsEnd, file_name, replicate) {
  dataSheet %>%
    filter(between(obs, obsStart, obsEnd)) %>%
    mutate(
      logNum   = obs - (obsStart - 1),
      cycle    = cycleAorB,
      treatment = cycleType,
      species  = spp,
      File     = file_name,
      IRGAName = IRGAName,
      replicate = replicate
    ) %>%
    select(date, IRGAName, species, cycle, treatment, replicate, logNum, A, gsw, Ci)
}

#taking all data from metadata/csvs, cleaning data, returning as one data frame
read_and_clean_all <- function(metadata, data_dir = "Data") {
  metadata <- metadata %>%
    mutate(File = as.character(File)) #taking file name, making String
  #For each row in metadata, read, clean, and return a tibble
  all_clean <- metadata %>%
    pmap_dfr(function(Date, Instrument, Species, Treatment, Cycle, Replicate, Start, End, File, Notes) {
      fullpath <- file.path(data_dir, File) # Build full file path
      raw <- readr::read_csv(fullpath, show_col_types = FALSE)  # Read raw CSV
      cleanData( #cleaning data to rename/only use needed cols
        dataSheet  = raw,
        IRGAName   = Instrument,  # use the metadata "Instrument" (PNNL/NoName/UT)
        cycleAorB  = Cycle,       # use the metadata "Cycle" (A/B, NA = blank)
        cycleType  = Treatment,   # use the metadata "Treatment" (Control, Fresh, Salt, etc)
        spp        = Species,     # use the metadata "Species" (ACRU, FAGR, LITU)
        obsStart   = Start,       # first log num
        obsEnd     = End,         # last log num
        file_name  = File,        # which .csv the data is from
        replicate  = Replicate    # use the metadata "Replicate"
      )
    })
  return(all_clean)
}

cleaned_all <- read_and_clean_all(metadata, data_dir = "Data")
glimpse(cleaned_all)


#Clean/manip data

#adjust based on blanks data
adjusted_all <- cleaned_all |> 
  group_by(IRGAName, treatment) |> 
    filter(treatment == "Blank") |> 
      summarize(offsetA = mean(A),
             offsetGSW = mean(gsw)) |>
  ungroup() |> 
  select(IRGAName, offsetA, offsetGSW) |> 
  full_join(cleaned_all, adjustmentFactors, by = "IRGAName") |> 
  mutate(adjA = A - offsetA,
         adjGSW = gsw - offsetGSW)
  

# #the one that would tell when/if it reaches x% decrease
placeholderName <- adjusted_all |> 
  group_by(species, treatment, cycle) %>%
  arrange(logNum) %>%
  mutate(aCheck = first(A) * 0.8, #0.8 = 20% decrease
         gswCheck = first(gsw) * 0.8) |> 
  mutate(isLessA = FALSE, isLessGSW = FALSE) #|> 
  # for((need to translate this, but i = 0; i < cycle.length, i++) {
  #   if(aCheck > A) {aCheck = TRUE}
  #   if(gswCheck > gsw) {gswCheck = TRUE}
  # })
  # for(i = 0; i < aCheck.length - 2, i++) {
  #     if(aCheck[i] == TRUE && aCheck[i+1] == TRUE && aCheck[i+2] == TRUE) {
  #       return i
  #     } +
  #     if(gswCheck[i] == TRUE && gswCheck[i+1] == TRUE && gswCheck[i+2] == TRUE){
  #       return i
  #     }
  # }
  
#   cleaned_all |> 
#     group_by(species, treatment, cycle, replicate) %>%
#     arrange(logNum) #|> 
#       #for the cycle, iterate one index at a time.
#       #if there are three consecutive "fail" at a time, return first index of fail
#     
#   #return logNum where first true if true, else return NA
# }


relative <- cleaned_all %>%
  group_by(species, treatment, cycle) %>%
  arrange(logNum) %>%
  mutate(A0 = first(A),
         gsw0 = first(gsw),
         A_rel = A/A0,
         gsw_rel = gsw/gsw0)
#%>% to calculate means of relative assmilation and stomatal conductance
# add to this summarize statement (can also do sd, counts, etc)
# summarize(mean_A = mean(A_rel))
#did this below!

relativeAdjusted <- adjusted_all |> 
  group_by(species, treatment, cycle) |> 
  arrange(logNum) |> 
  mutate(A0 = first(adjA),
         gsw0 = first(adjA),
         A_rel = adjA/A0,
         gsw_rel = adjGSW/gsw0) |> 
  ungroup() |> 
  filter(!is.na(cycle), A < 12) |> 
  group_by(species, logNum, treatment, cycle) |> 
  summarize(A_mean = mean(A_rel),
            A_sd = sd(A_rel),
            gsw_mean = mean(gsw_rel),
            gsw_sd = sd(gsw_rel))

```

# Graphs
ideas: 
can i color code things more properly?
how to make graphs more readable?

## with K: 
```{r all data graphs, include = FALSE}
###
### K chunk from August 12th meeting
###

#cleaning data step first (so we only have to do it once)
data <- cleaned_all %>%
  filter(! is.na(cleaned_all$cycle),
         A < 12) #remove outliers in A
#(i like having the new var name out front to easily find it later!)

#making means data frame for plotting
clean_means <- data %>%
  group_by(species, logNum, treatment, cycle) %>%
  summarize(A_mean = mean(A),
            A_sd = sd(A),
            gsw_mean = mean(gsw),
            gsw_sd = sd(gsw)) #-> clean_means

#making rolled time series data frame for plotting
#rolled average = plots the average of this and the next two (n=3), smooths curve
rolled <- data %>%
  group_by(species, treatment, cycle, replicate) %>%
  arrange(logNum) %>%
  mutate(A_roll = zoo::rollmean(A, k = 3, align = "left", fill = NA),
         gsw_roll = zoo::rollmean(gsw, k = 3, align = "left", fill = NA)) # -> rolled

#and lastly, making an average of the rolled values for plotting
#may decide later that any or all of these are not useful
rolled_means <- rolled %>%
  ungroup() %>%
  group_by(species, logNum, treatment, cycle) %>%
  summarize(A_mean = mean(A_roll),
            A_sd = sd(A_roll),
            gsw_mean = mean(gsw_roll),
            gsw_sd = sd(gsw_roll)) #-> rolled_means


#a half-decent plot with errorbars
ggplot() + geom_line(data = rolled_means, aes(logNum, A_mean, color = treatment),
                     linewidth = 1.25) + 
  geom_point(data = clean_means, aes(logNum, A_mean, color = treatment), alpha = 0.5) + 
  geom_errorbar(data = clean_means, aes(x = logNum, ymin = A_mean - A_sd, ymax = A_mean + A_sd,
                                       color = treatment), alpha = 0.3) +
  geom_hline(yintercept = 0, color = 'red') + #adds red line on 0
  facet_grid(species ~ cycle, scales = "free_y") + theme_bw()

#attempt to graph with ribbons instead of lines
ggplot() + geom_line(data = rolled_means, aes(logNum, A_mean, color = treatment),
                     linewidth = 1.25) + 
  geom_point(data = clean_means, aes(logNum, A_mean, color = treatment), alpha = 0.5) + 
  geom_ribbon(data = clean_means, aes(x = logNum, ymin = A_mean - A_sd, ymax = A_mean + A_sd,
                                       color = treatment, fill = treatment), alpha = 0.1) +
  geom_hline(yintercept = 0, color = 'red') +
  facet_grid(species ~ cycle, scales = "free_y") + theme_bw()


#same as above, but with gsw
ggplot() + geom_line(data = rolled_means, aes(logNum, gsw_mean, color = treatment),
                     linewidth = 1.25) + 
  geom_point(data = clean_means, aes(logNum, gsw_mean, color = treatment), alpha = 0.5) + 
  geom_errorbar(data = clean_means, aes(x = logNum, ymin = gsw_mean - gsw_sd, ymax = gsw_mean + gsw_sd,
                                       color = treatment), alpha = 0.3) +
  geom_hline(yintercept = 0, color = 'red') + #adds red line on 0
  facet_grid(species ~ cycle, scales = "free_y") + theme_bw()

ggplot() + geom_line(data = rolled_means, aes(logNum, gsw_mean, color = treatment),
                     linewidth = 1.25) + 
  geom_point(data = clean_means, aes(logNum, gsw_mean, color = treatment), alpha = 0.5) + 
  #geom_ribbon(data = clean_means, aes(x = logNum, ymin = gsw_mean - gsw_sd, ymax = gsw_mean + gsw_sd,
                                      # color = treatment, fill = treatment), alpha = 0.3) +
  geom_hline(yintercept = 0, color = 'red') + #adds red line on 0
  facet_grid(species ~ cycle)  + theme_bw()


#K skipped this
# cleaned_all %>%
#  filter(!is.na(cycle),
#         replicate = 1) %>%
# ggplot(aes(logNum, gsw, color = treatment)) +
#   geom_point() + geom_line() + facet_grid(species ~ cycle, scales = "free_y") +
#   ggtitle("Example Stomatal Conductance Panel")



#NEEDS TO ADJUST: 
##can i color-code the lines? (detached = red, control = green, bucket = cyan?, salt = yellow, fresh = blue)



#Prior graphs
#make sure to rename with adj variable data set
ggplot(data = subset(cleaned_all, !is.na(cycle)), aes(logNum, A, color = treatment)) +
  geom_point(aes(alpha = 0.5)) +
  geom_line(aes(alpha = 0.5)) + facet_grid(species ~ cycle, scales = "free_y") + #~ creates the grid row/col
  ggtitle("Example Assimilation Panel")

ggplot(subset(cleaned_all, !is.na(cycle)), aes(logNum, gsw, color = treatment)) +
  geom_point() + geom_line() + facet_grid(species ~ cycle, scales = "free_y") +
  ggtitle("Example Stomatal Conductance Panel")

#these used "relative" data set from the original document
ggplot(subset(relative,!is.na(cycle)), aes(logNum, A_rel, color = treatment)) +
  geom_point() + geom_line() + facet_grid(species ~ cycle) +
  ggtitle("Example Assimilation Panel")

ggplot(subset(relative,!is.na(cycle)), aes(logNum, gsw_rel, color = treatment)) +
  geom_point() + geom_line() + facet_grid(species ~ cycle, scales = "free_y") +
  ggtitle("Example Stomatal Conductance Panel")
```

```{r all data graphs but with adjusted values (it broke the first time), include = FALSE}
###
### K chunk from August 12th meeting
###

#cleaning data step first (so we only have to do it once)
data <- adjusted_all %>%
  filter(!is.na(adjusted_all$cycle),
         A < 12) #remove outliers in A
#(i like having the new var name out front to easily find it later!)

#making means data frame for plotting
#this is making a mean/stdev for each cycle log 
##aka averaging all ACRU fresh log 1s
clean_means <- data %>%
  group_by(species, logNum, treatment, cycle) %>%
  summarize(A_mean = mean(adjA),
            A_sd = sd(adjA),
            gsw_mean = mean(adjGSW),
            gsw_sd = sd(adjGSW)) |> 
  ungroup() #-> clean_means
#does this ever get called??? <- no, it was a remnant from first graph attempt


#making rolled time series data frame for plotting
#rolled average = plots the average of this and the next two (n=3), smooths curve
rolled <- data %>%
  group_by(species, treatment, cycle, replicate) %>%
  arrange(logNum) %>%
  mutate(A_roll = rollmean(adjA, k = 3, align = "left", fill = NA),
         gsw_roll = rollmean(adjGSW, k = 3, align = "left", fill = NA)) # -> rolled

#and lastly, making an average of the rolled values for plotting
#may decide later that any or all of these are not useful
rolled_means <- rolled %>%
  ungroup() %>%
  group_by(species, logNum, treatment, cycle) %>%
  summarize(A_mean = mean(A_roll),
            A_sd = sd(A_roll),
            gsw_mean = mean(gsw_roll),
            gsw_sd = sd(gsw_roll)) #-> rolled_means


#a half-decent plot with errorbars
ggplot() + geom_line(data = rolled_means, aes(logNum, A_mean, color = treatment),
                     linewidth = 1.25) + 
  geom_point(data = clean_means, aes(logNum, A_mean, color = treatment), alpha = 0.5) + 
  geom_errorbar(data = clean_means, aes(x = logNum, ymin = A_mean - A_sd, ymax = A_mean + A_sd,
                                       color = treatment), alpha = 0.3) +
  geom_hline(yintercept = 0, color = 'red') + #adds red line on 0
  facet_grid(species ~ cycle, scales = "free_y") + theme_bw()

#attempt to graph with ribbons instead of lines
#i don't really like this one- too much going on!
ggplot() + geom_line(data = rolled_means, aes(logNum, A_mean, color = treatment),
                     linewidth = 1.25) + 
  geom_point(data = clean_means, aes(logNum, A_mean, color = treatment), alpha = 0.5) + 
  geom_ribbon(data = clean_means, aes(x = logNum, ymin = A_mean - A_sd, ymax = A_mean + A_sd,
                                       color = treatment, fill = treatment), alpha = 0.2, color = NA) +
  geom_hline(yintercept = 0, color = 'red') +
  facet_grid(species ~ cycle, scales = "free_y") + theme_bw()


#same as above, but with gsw
ggplot() + geom_line(data = rolled_means, aes(logNum, gsw_mean, color = treatment),
                     linewidth = 1.25) + 
  geom_point(data = clean_means, aes(logNum, gsw_mean, color = treatment), alpha = 0.5) + 
  geom_errorbar(data = clean_means, aes(x = logNum, ymin = gsw_mean - gsw_sd, ymax = gsw_mean + gsw_sd,
                                       color = treatment), alpha = 0.3) +
  geom_hline(yintercept = 0, color = 'red') + #adds red line on 0
  facet_grid(species ~ cycle, scales = "free_y") + theme_bw()

ggplot() + geom_line(data = rolled_means, aes(logNum, gsw_mean, color = treatment),
                     linewidth = 1.25) + 
  geom_point(data = clean_means, aes(logNum, gsw_mean, color = treatment), alpha = 0.5) + 
  geom_ribbon(data = clean_means, aes(x = logNum, ymin = gsw_mean - gsw_sd, ymax = gsw_mean + gsw_sd,
                                      color = treatment, fill = treatment), alpha = 0.3, color = NA) +
  geom_hline(yintercept = 0, color = 'red') + #adds red line on 0
  facet_grid(species ~ cycle)  + theme_bw()


#K skipped this
# cleaned_all %>%
#  filter(!is.na(cycle),
#         replicate = 1) %>%
# ggplot(aes(logNum, gsw, color = treatment)) +
#   geom_point() + geom_line() + facet_grid(species ~ cycle, scales = "free_y") +
#   ggtitle("Example Stomatal Conductance Panel")



#NEEDS TO ADJUST: 
##can i color-code the lines? (detached = red, control = green, bucket = cyan?, salt = yellow, fresh = blue)



#Prior graphs
#make sure to rename with adj variable data set
ggplot(data = subset(cleaned_all, !is.na(cycle)), aes(logNum, A, color = treatment)) +
  geom_point(aes(alpha = 0.5)) +
  geom_line(aes(alpha = 0.5)) + facet_grid(species ~ cycle, scales = "free_y") + #~ creates the grid row/col
  ggtitle("Example Assimilation Panel")

ggplot(subset(cleaned_all, !is.na(cycle)), aes(logNum, gsw, color = treatment)) +
  geom_point() + geom_line() + facet_grid(species ~ cycle, scales = "free_y") +
  ggtitle("Example Stomatal Conductance Panel")

#these used "relative" data set from the original document
ggplot(subset(relative,!is.na(cycle)), aes(logNum, A_rel, color = treatment)) +
  geom_point() + geom_line() + facet_grid(species ~ cycle) +
  ggtitle("Example Assimilation Panel")

ggplot(subset(relative,!is.na(cycle)), aes(logNum, gsw_rel, color = treatment)) +
  geom_point() + geom_line() + facet_grid(species ~ cycle, scales = "free_y") +
  ggtitle("Example Stomatal Conductance Panel")
```


#Statistics
```{r Statistics, echo = FALSE}

#a simple test of bucket effect

#isolate data for testing effect
simple_bucket_test <-  data %>%
  ungroup() %>%
  group_by(treatment, species, replicate) %>%
  filter(cycle == "A",
         treatment != "Detached",
         logNum < 26 & logNum > 14) %>%
  summarize(A_mean = mean(A),
            gsw_mean = mean(gsw)) #-> simple_bucket_test

#Effect on photosynthetic rate
simple_bucket1 <- aov(A_mean ~ treatment * species,
                    data = simple_bucket_test) 
summary(simple_bucket1)


#Effect on stomatal conductance
simple_bucket2 <- aov(gsw_mean ~ treatment * species,
                    data = simple_bucket_test) 
summary(simple_bucket2)



#combining cylces to double the statistics
combinedCyclesData <- data |> 
  ungroup() |> 
  mutate(across("treatment", str_replace, "Fresh", "Bucket")) |> 
  group_by(treatment, species, replicate) |> 
  filter(treatment != "Detached", treatment != "Salt",
         logNum < 26 & logNum > 14) |> 
  summarize(A_mean = mean(A),
            gsw_mean = mean(gsw))

simpleDoubleBucketA <- aov(A_mean ~ treatment*species, 
                           data = combinedCyclesData)
summary(simpleDoubleBucketA)

simpleDoubleBucketGSW <- aov(A_mean ~ treatment * species,
                             data = combinedCyclesData)
summary(simpleDoubleBucketGSW)


```


## by species
``` {r graph by species NEW, echo = FALSE, include = fALSE} 




```

#old work
##old import/separating
```{r import data OLD, include = FALSE}
UTDay1 <- read.csv("2026-07-21_UT.csv") #LITU1 detached, LITU1 salt, FAGR1 detached
NoNameDay1 <- read.csv("2026-07-21_NoName.csv") #LITU1 bucket, LITU1 fresh, FAGR1 bucket
PNNLDay1 <- read.csv("2026-07-21 PNNL Day 1.csv") #LITU1 control A, LITU1 control B, FAGR1 control A

UTDay2 <- read.csv("2026-07-22 UT Day 2.csv") #LITU2 detached, LITU2 salt, FAGR1 salt, blank = obs 
NoNameDay2 <- read.csv("2026-07-22 NoName Day 2.csv") #LITU2 bucket, LITU2 fresh, FAGR1 fresh, blank = obs 115
PNNLDay2 <- read.csv("2026-07-22 PNNL Day 2.csv") #LITU2 controlA, LITU2 controlB, FAGR2 controlB, blank = obs 113

UTDay3 <- read.csv("2026-07-24 UT Day 3.csv") #FAGR2 detached, salt, blank = 77
NoNameDay3 <- read.csv("2026-07-24 NoName (misnamed PNNL)Day 3.csv") #FAGR2 control A, control b, blank = NA
PNNLDay3 <- read.csv("2026-07-24 PNNL (misnamed NoName) Day 3.csv") #FAGR2 bucket, fresh, blank = 77

UTDay4 <- read.csv("2026-07-29 UT Day 4.csv") #FAGR3 detached, FAGR3 salt, LITU3 salt
NoNameDay4 <- read.csv("2026-07-29 NoName Day 4.csv") #FAGR3 bucket, fresh, LITU3 fresh
PNNLDay4 <- read.csv("2026-07-29 PNNL Day 4.csv") #FAGR3 controlA,B, LITU3 controlB

NoNameDay5 <- read.csv("2026-07-30 noname day 5.csv") #LITU3 controlA
UTDay5 <- read.csv("2026-07-30 ut day 5.csv") #LITU3 bucket
PNNLDay5 <- read.csv("2026-07-30 pnnl day 5.csv") #LITU3 detached

UTDay7 <- read.csv("data/2026-08-10 UT Day 7.csv") #controls
NoNameDay7 <- read.csv("data/2026-08-10 NoName Day 7.csv") #bucket fresh bucket
PNNLDay7 <- read.csv("data/2026-08-10 PNNL Day 7.csv") #detached salt detached

UTDay8 <- read.csv("data/2026-08-11 UT Day 8.csv") #controls
NoNameDay8 <- read.csv("data/2026-08-11 NoName Day 8.csv") #fresh and bucket
PNNLDay8 <- read.csv("data/2026-08-11 PNNL Day 8.csv")#salt detached salt
```
```{r seperate into cycles OLD, include = FALSE}
#could i make this a function?? does r do functions???
#I MADE IT A FUNCTION!!! (it works!!) YAY!!
cleanData <- function(dataSheet, cycleNumber, cycleType, spp, obsStart, obsEnd) {
  #dataSheet = data, cylceNumber = String, cycleType = String, spp = String, obsStart = int, obsEnd = int
  dataSheet %>% #grabs from correct data sheet
    filter(between(obs, obsStart, obsEnd)) %>% #filters to correct log numbers
    mutate(logNum = obs - (obsStart- 1)) %>% #makes it so logNum always starts on 1
    mutate(name = cycleNumber) %>% #(like LITU_A1 = tulip cycle a rep 1), to ID cycle #
    mutate(cycle = cycleType) %>% #adds cycle type
    mutate(species = spp) %>% #adds species
    select(logNum, A, gsw, date, Ci, cycle, species) #get data!
}
#metadata sheet?

#LITU
LITU_A1_control <- cleanData(PNNLDay1, "LITU_A1", "controlA", "LITU", 2, 39)
LITU_A1_detached <- cleanData(UTDay1, "LITU_A1", "detached", "LITU", 1, 38)
LITU_A1_bucket <- cleanData(NoNameDay1, "LITU_A1", "bucket", "LITU", 1, 38)

LITU_B1_control <- cleanData(PNNLDay1, "LITU_B1", "controlB", "LITU", 41, 77)
LITU_B1_fresh <- cleanData(NoNameDay1, "LITU_B1", "fresh", "LITU", 40, 77)
LITU_B1_salt <- cleanData(UTDay1, "LITU_B1", "salt", "LITU", 40, 77)

LITU_A2_control <- cleanData(PNNLDay2, "LITU_A2", "controlA", "LITU", 1, 38)
LITU_A2_detached <- cleanData(UTDay2, "LITU_A2", "detached", "LITU", 1, 38)
LITU_A2_bucket <- cleanData(NoNameDay2, "LITU_A2", "bucket", "LITU", 1, 38)

LITU_B2_control <- cleanData(PNNLDay2, "LITU_B2", "controlB", "LITU", 39, 75)
LITU_B2_fresh <- cleanData(NoNameDay2, "LITU_B2", "fresh", "LITU", 39, 76)
LITU_B2_salt <- cleanData(UTDay2, "LITU_B2", "salt", "LITU", 39, 76)

#need to add in both LITU remaining cycles

LITU_A3_control <- cleanData(NoNameDay5, "LITU_A3", "controlA", "LITU", 1, 37)
LITU_A3_detached <- cleanData(PNNLDay5, "LITU_A3", "detached", "LITU", 1, 38)
LITU_A3_bucket <- cleanData(UTDay5, "LITU_A3", "bucket", "LITU", 1, 38)

LITU_B3_control <- cleanData(PNNLDay4, "LITU_B3", "controlB", "LITU", 75, 111)
LITU_B3_fresh <- cleanData(NoNameDay4, "LITU_B3", "fresh", "LITU", 77, 114)
LITU_B3_salt <- cleanData(UTDay4, "LITU_B3", "salt", "LITU", 77, 114)


#FAGR
FAGR_A1_control <- cleanData(PNNLDay1, "FAGR_A1", "controlA", "FAGR", 79, 116)
FAGR_A1_detached <- cleanData(UTDay1, "FAGR_A1", "detached", "FAGR", 79, 116)
FAGR_A1_bucket <- cleanData(NoNameDay1, "FAGR_A1", "bucket", "FAGR", 80, 117)

FAGR_B1_control <- cleanData(PNNLDay2, "FAGR_B1", "controlB", "FAGR", 76, 112)
FAGR_B1_fresh <- cleanData(NoNameDay2, "FAGR_B1", "fresh", "FAGR", 77, 114)
FAGR_B1_salt <- cleanData(UTDay2, "FAGR_B1", "salt", "FAGR", 77, 114)

FAGR_A2_control <- cleanData(NoNameDay3, "FAGR_A2", "controlA", "FAGR", 1, 37)
FAGR_A2_detached <- cleanData(UTDay3, "FAGR_A2", "detached", "FAGR", 1, 38)
FAGR_A2_bucket <- cleanData(PNNLDay3, "FAGR_A2", "bucket", "FAGR", 1, 38)

FAGR_B2_control <- cleanData(NoNameDay3, "FAGR_B2", "controlB", "FAGR", 38, 74)
FAGR_B2_fresh <- cleanData(PNNLDay3, "FAGR_B2", "fresh", "FAGR", 39, 76)
FAGR_B2_salt <- cleanData(UTDay3, "FAGR_B2", "salt", "FAGR", 39, 76)

FAGR_A3_control <- cleanData(NoNameDay4, "FAGR_A3", "controlA", "FAGR", 1, 37)
FAGR_A3_detached <- cleanData(UTDay4, "FAGR_A3", "detached", "FAGR", 1, 38)
FAGR_A3_bucket <- cleanData(PNNLDay4, "FAGR_A3", "bucket", "FAGR", 1, 38)

FAGR_B3_control <- cleanData(NoNameDay4, "FAGR_B3", "controlB", "FAGR", 38, 74)
FAGR_B3_fresh <- cleanData(PNNLDay4, "FAGR_B3", "fresh", "FAGR", 39, 76)
FAGR_B3_salt <- cleanData(UTDay4, "FAGR_B3", "salt", "FAGR", 39, 76)



ACRU_A1_control <- cleanData(UTDay7, "ACRU_A1", "controlA", "ACRU", 1, 38)
ACRU_A1_detached <- cleanData(PNNLDay7, "ACRU_A1", "detached", "ACRU", 1, 38)
ACRU_A1_bucket <- cleanData(NoNameDay7, "ACRU_A1", "bucket", "ACRU", 1, 38)

ACRU_B1_control <- cleanData(UTDay7, "ACRU_B1", "controlB", "ACRU", 39, 76)
ACRU_B1_fresh <- cleanData(NoNameDay7, "ACRU_B1", "fresh", "ACRU", 39, 76)
ACRU_B1_salt <- cleanData(PNNLDay7, "ACRU_B1", "salt", "ACRU", 39, 76)

ACRU_A2_control <- cleanData(UTDay7, "ACRU_A1", "controlA", "ACRU", 77, 114)
ACRU_A2_detached <- cleanData(PNNLDay7, "ACRU_A1", "detached", "ACRU", 77, 114)
ACRU_A2_bucket <- cleanData(NoNameDay7, "ACRU_A1", "bucket", "ACRU", 77, 114)

ACRU_B2_control <- cleanData(UTDay8, "ACRU_B2", "controlB", "ACRU", 1,38)
ACRU_B2_fresh <- cleanData(NoNameDay8, "ACRU_B2", "fresh", "ACRU", 1,38)
ACRU_B2_salt <- cleanData(PNNLDay8, "ACRU_B2", "salt", "ACRU", 1,38)

ACRU_A3_control <- cleanData(UTDay8, "ACRU_A3", "controlA", "ACRU", 39, 76)
ACRU_A3_detached <- cleanData(PNNLDay8, "ACRU_A3", "detached", "ACRU", 39, 76)
ACRU_A3_bucket <- cleanData(NoNameDay8, "ACRU_A3", "bucket", "ACRU", 39, 76)

ACRU_B3_control <- cleanData(UTDay8, "ACRU_B3", "controlB", "ACRU", 77,114)
ACRU_B3_salt <- cleanData(PNNLDay8, "ACRU_B3", "salt", "ACRU", 77,114


```

## old cycle graphing: 
```{r graph one cycle, echo=FALSE}

##Functions ----
plotOneCycleA <- function(data1, data2, data3, title = "No title provided", lowerBound = -1, upperBound = 10) {
  #data1,2,3 = dataset, title = String, bounds = num (default 0:4)
  ggplot() + 
    geom_line(data = data1, aes(x = logNum, y = A, color = cycle)) +
    geom_line(data = data2, aes(x = logNum, y = A, color = cycle)) +
    geom_line(data = data3, aes(x = logNum, y = A, color = cycle)) +
    geom_point(data = data1, aes(x = logNum, y = A, color = cycle)) +
    geom_point(data = data2, aes(x = logNum, y = A, color = cycle)) +
    geom_point(data = data3, aes(x = logNum, y = A, color = cycle)) +
    theme_bw() +
    ggtitle(title) + 
    coord_cartesian(ylim = c(lowerBound, upperBound))
} 

plotOneCycleGSW <- function(data1, data2, data3, title = "No title provided", lowerBound = 0, upperBound = 0.3) { 
  #NOT WOKRING- WHY?
  #data1,2,3 = dataset, title = String, bounds = num (default -1:1)
  ggplot() + 
    geom_line(data = data1, aes(x = logNum, y = gsw, color = cycle)) +
    geom_line(data = data2, aes(x = logNum, y = gsw, color = cycle)) +
    geom_line(data = data3, aes(x = logNum, y = gsw, color = cycle)) +
    geom_point(data = data1, aes(x = logNum, y = gsw, color = cycle)) +
    geom_point(data = data2, aes(x = logNum, y = gsw, color = cycle)) +
    geom_point(data = data3, aes(x = logNum, y = gsw, color = cycle)) +
    theme_bw() +
    ggtitle(title) + 
    coord_cartesian(ylim = c(lowerBound, upperBound))
} 
#Debug: should make it so that cycle color is consistent regardless of input order. how to?

#LITU----
LITU_A1_graphA <- plotOneCycleA(LITU_A1_control, LITU_A1_bucket, LITU_A1_detached, "LITU A1", 0, 8)
LITU_A1_graphA
LITU_A1_graphGSW <- plotOneCycleGSW(LITU_A1_bucket, LITU_A1_control, LITU_A1_detached, "LITU A1", 0, 0.2)
LITU_A1_graphGSW

LITU_B1_graphA <- plotOneCycleA(LITU_B1_control, LITU_B1_fresh, LITU_B1_salt, "LITU B1", 0, 8)
LITU_B1_graphA
LITU_B1_graphGSW <- plotOneCycleGSW(LITU_B1_control, LITU_B1_fresh, LITU_B1_salt, "LITU B1", 0, 0.3)
LITU_B1_graphGSW

LITU_A2_graphA <- plotOneCycleA(LITU_A2_control, LITU_A2_bucket, LITU_A2_detached, "LITU A2", -1, 8)
LITU_A2_graphA
LITU_A2_graphGSW <- plotOneCycleGSW(LITU_A2_control, LITU_A2_bucket, LITU_A2_detached, "LITU A2", 0, 0.3)
LITU_A2_graphGSW

LITU_B2_graphA <- plotOneCycleA(LITU_B2_control, LITU_B2_fresh, LITU_B2_salt, "LITU B2", -1, 8)
LITU_B2_graphA
LITU_B2_graphGSW <- plotOneCycleGSW(LITU_B2_control, LITU_B2_fresh, LITU_B2_salt, "LITU B2", 0, 0.3)
LITU_B2_graphGSW

LITU_A3_graphA <- plotOneCycleA(LITU_A3_control, LITU_A3_detached, LITU_A3_bucket, "LITU A3", -1, 8)
LITU_A3_graphA
LITU_A3_graphGSW <- plotOneCycleGSW(LITU_A3_control, LITU_A3_detached, LITU_A3_bucket, "LITU A3", 0, 0.3)
LITU_A3_graphGSW

LITU_B3_graphA <- plotOneCycleA(LITU_B3_control, LITU_B3_fresh, LITU_B3_salt, "LITU B3", -1, 8)
LITU_B3_graphA
LITU_B3_graphGSW <- plotOneCycleGSW(LITU_B3_control, LITU_B3_fresh, LITU_B3_salt, "LITU B3", 0, 0.1)
LITU_B3_graphGSW

#FAGR ----
FAGR_A1_graphA <- plotOneCycleA(FAGR_A1_bucket, FAGR_A1_control, FAGR_A1_detached, "FAGR A1", -1, 7)
FAGR_A1_graphA
FAGR_A1_graphGSW <- plotOneCycleGSW(FAGR_A1_bucket, FAGR_A1_control, FAGR_A1_detached, "FAGR A1", 0, 0.05)
FAGR_A1_graphGSW

FAGR_B1_graphA <- plotOneCycleA(FAGR_B1_control, FAGR_B1_fresh, FAGR_B1_salt, "FAGR B1", -1, 7)
FAGR_B1_graphA
FAGR_B1_graphGSW <- plotOneCycleGSW(FAGR_B1_control, FAGR_B1_fresh, FAGR_B1_salt, "FAGR B1", 0, 0.1)
FAGR_B1_graphGSW

FAGR_A2_graphA <- plotOneCycleA(FAGR_A2_bucket, FAGR_A2_control, FAGR_A2_detached, "FAGR A2", -1, 5)
FAGR_A2_graphA 
FAGR_A2_graphGSW <- plotOneCycleGSW(FAGR_A2_bucket, FAGR_A2_control, FAGR_A2_detached, "FAGR A2", 0, 0.1)
FAGR_A2_graphGSW

FAGR_B2_graphA <- plotOneCycleA(FAGR_B2_control, FAGR_B2_fresh, FAGR_B2_salt, "FAGR B2", -1, 5)
FAGR_B2_graphA
FAGR_B2_graphGSW <- plotOneCycleGSW(FAGR_B2_control, FAGR_B2_fresh, FAGR_B2_salt, "FAGR B2", 0, 0.1)
FAGR_B2_graphGSW

FAGR_A3_graphA <- plotOneCycleA(FAGR_A3_bucket, FAGR_A3_control, FAGR_A3_detached, "FAGR A3", 0, 6)
FAGR_A3_graphA 
FAGR_A3_graphGSW <- plotOneCycleGSW(FAGR_A3_bucket, FAGR_A3_control, FAGR_A3_detached, "FAGR A3", 0, 0.1)
FAGR_A3_graphGSW

FAGR_B3_graphA <- plotOneCycleA(FAGR_B3_control, FAGR_B3_fresh, FAGR_B3_salt, "FAGR B3", 0, 6)
FAGR_B3_graphA
FAGR_B3_graphGSW <- plotOneCycleGSW(FAGR_B3_control, FAGR_B3_fresh, FAGR_B3_salt, "FAGR B3", 0, 0.1)
FAGR_B3_graphGSW

#ACRU ----
ACRU_A1_graphA <- plotOneCycleA(ACRU_B1_fresh, ACRU_B1_control, ACRU_B1_salt, "ACRU A1", -1, 7)
ACRU_A1_graphA
ACRU_A1_graphGSW <- plotOneCycleGSW(ACRU_B1_fresh, ACRU_B1_control, ACRU_B1_salt, "ACRU A1", -0.1, 0.1)
ACRU_A1_graphGSW

ACRU_B1_graphA <- plotOneCycleA(ACRU_A1_bucket, ACRU_A1_control, ACRU_A1_detached, "ACRU A1", -1, 7)
ACRU_B1_graphA
ACRU_B1_graphGSW <- plotOneCycleGSW(ACRU_A1_bucket, ACRU_A1_control, ACRU_A1_detached, "ACRU A1", -0.1, 0.1)
ACRU_B1_graphGSW

#----
#from original code
#ggplot() + 
 # geom_line(data = control2, aes(x = logNum, y = A, color = cycle)) +
  #geom_line(data = fresh, aes(x = logNum, y = A, color = cycle)) +
 # geom_line(data = salt, aes(x = logNum, y = A, color = cycle))+ 
 # theme_bw() + 
 # ggtitle("Cycle 2 (buckets)") + 
 # coord_cartesian(ylim = c(0, 4))
```


## old summary graphing
``` {r graph by species test OLD, echo = FALSE, include = FALSE}

#need to combine species into one variable
  # c can smash them together, but doesn't really help?
#how to? 
  #new variable called [species][treatment]Combined[A/gsw]
  #maybe it takes [col] from [data frame], adds together and /3 for each value in the row?
  #no, needs to average! (some control's don't have 3 obs bc i got lazy)


#THIS IS VERY MESSY. CLEAN LATER


#new idea: make new var ([tree]_[treatment]) with obs column (1:38), A and gsw for each replicate, and the avg of the three

#i was having trouble with pipe logic so i just made this a separate function
mergeData <- function(data1, data2, data3) {
  halfMerge = merge(data1, data2, by = "logNum", all.x = TRUE) 
  merge(halfMerge, data3, by = "logNum", all.x = TRUE)
}

#THSI DOES NOT WORK
combineData <- function(data1, data2, data3) {
   data = mergeData(data1, data2, data3) %>%  
   rename(A1 = A.x) %>% 
   rename(A2 = A.y) %>% 
   rename(A3 = A) %>% 
   rename(gsw1 = gsw.x) %>% 
   rename(gsw2 = gsw.y) %>% 
   rename(gsw3 = gsw) %>% 
   select(logNum, A1, A2, A3, gsw1, gsw2, gsw3, species, cycle) 
   #get the averages here
   mutate(data, avgA = rowMeans(select(data, c(A1, A2, A3)), na.rm = TRUE)) %>%  
   mutate(data, avgGSW = rowMeans(select(data, c(gsw1, gsw2, gsw3)), na.rm = TRUE))
}

LITUcontrolA <- combineData(LITU_A1_control, LITU_A2_control, LITU_A3_control)
LITUbucket <- combineData(LITU_A1_bucket, LITU_A2_bucket, LITU_A3_bucket)
LITUdetached <- combineData(LITU_A1_detached, LITU_A2_detached, LITU_A3_detached)
LITUcontrolB <- combineData(LITU_B1_control, LITU_B2_control, LITU_B3_control)
LITUsalt <- combineData(LITU_B1_salt, LITU_B2_salt, LITU_B3_salt)
LITUfresh <- combineData(LITU_B1_fresh, LITU_B2_fresh, LITU_B3_fresh)

FAGRcontrolA <- combineData(FAGR_A1_control, FAGR_A2_control, FAGR_A3_control)
FAGRbucket <- combineData(FAGR_A1_bucket, FAGR_A2_bucket, FAGR_A3_bucket)
FAGRdetached <- combineData(FAGR_A1_detached, FAGR_A2_detached, FAGR_A3_detached)
FAGRcontrolB <- combineData(FAGR_B1_control, FAGR_B2_control, FAGR_B3_control)
FAGRsalt <- combineData(FAGR_B1_salt, FAGR_B2_salt, FAGR_B3_salt)
FAGRfresh <- combineData(FAGR_B1_fresh, FAGR_B2_fresh, FAGR_B3_fresh)


plotAvgA <- function(data1, data2, data3, title = "No title provided", lowerBound = -1, upperBound = 10) {
  #data1,2,3 = dataset, title = String, bounds = num (default 0:4)
  ggplot() + 
    geom_line(data = subset(data1, !is.na(avgA)), aes(x = logNum, y = avgA, color = cycle)) +
    geom_line(data = subset(data2, !is.na(avgA)), aes(x = logNum, y = avgA, color = cycle)) +
    geom_line(data = subset(data3, !is.na(avgA)), aes(x = logNum, y = avgA, color = cycle)) +
    geom_point(data = subset(data1, !is.na(avgA)), aes(x = logNum, y = avgA, color = cycle)) +
    geom_point(data = subset(data2, !is.na(avgA)), aes(x = logNum, y = avgA, color = cycle)) +
    geom_point(data = subset(data3, !is.na(avgA)), aes(x = logNum, y = avgA, color = cycle)) +
    theme_bw() +
    ggtitle(title) + 
    coord_cartesian(ylim = c(lowerBound, upperBound))
}

plotAvgGSW <- function(data1, data2, data3, title = "No title provided", lowerBound = 0, upperBound = 0.2) {
  #data1,2,3 = dataset, title = String, bounds = num (default 0:4)
  ggplot() + 
    geom_line(data = subset(data1, !is.na(avgA)), aes(x = logNum, y = avgGSW, color = cycle)) +
    geom_line(data = subset(data2, !is.na(avgA)), aes(x = logNum, y = avgGSW, color = cycle)) +
    geom_line(data = subset(data3, !is.na(avgA)), aes(x = logNum, y = avgGSW, color = cycle)) +
    geom_point(data = subset(data1, !is.na(avgA)), aes(x = logNum, y = avgGSW, color = cycle)) +
    geom_point(data = subset(data2, !is.na(avgA)), aes(x = logNum, y = avgGSW, color = cycle)) +
    geom_point(data = subset(data3, !is.na(avgA)), aes(x = logNum, y = avgGSW, color = cycle)) +
    theme_bw() +
    ggtitle(title) + 
    coord_cartesian(ylim = c(lowerBound, upperBound))
}


#clean this up later
LITU_A_avgA <- plotAvgA(LITUcontrolA, LITUbucket, LITUdetached)
LITU_A_avgA
LITU_B_avgA <- plotAvgA(LITUcontrolB, LITUfresh, LITUsalt)
LITU_B_avgA

LITU_A_avgGSW <- plotAvgGSW(LITUcontrolA, LITUbucket, LITUdetached, "LITU A gsw", 0, 0.2)
LITU_A_avgGSW
LITU_B_avgGSW <- plotAvgGSW(LITUcontrolB, LITUfresh, LITUsalt, "LITU B gsw", 0, 0.2)
LITU_B_avgGSW

ggplot() + 
  geom_line(data = LITUcontrolA, aes(x = logNum, y = avgA, color = cycle)) +
  geom_line(data = LITUcontrolB, aes(x = logNum, y = avgA, color = cycle)) + 
  geom_line(data = LITUbucket, aes(x = logNum, y = avgA, color = cycle)) +
  geom_line(data = LITUdetached, aes(x = logNum, y = avgA, color = cycle)) +
  geom_line(data = LITUfresh, aes(x = logNum, y = avgA, color = cycle)) + 
  geom_line(data = LITUsalt, aes(x = logNum, y = avgA, color = cycle)) +
  geom_point(data = LITUcontrolA, aes(x = logNum, y = avgA, color = cycle)) +
  geom_point(data = LITUcontrolB, aes(x = logNum, y = avgA, color = cycle)) + 
  geom_point(data = LITUbucket, aes(x = logNum, y = avgA, color = cycle)) +
  geom_point(data = LITUdetached, aes(x = logNum, y = avgA, color = cycle)) +
  geom_point(data = LITUfresh, aes(x = logNum, y = avgA, color = cycle)) + 
  geom_point(data = LITUsalt, aes(x = logNum, y = avgA, color = cycle)) +
  theme_bw() +
  ggtitle("all together litu") 

ggplot() + 
  geom_line(data = LITUcontrolA, aes(x = logNum, y = avgGSW, color = cycle)) +
  geom_line(data = LITUcontrolB, aes(x = logNum, y = avgGSW, color = cycle)) + 
  geom_line(data = LITUbucket, aes(x = logNum, y = avgGSW, color = cycle)) +
  geom_line(data = LITUdetached, aes(x = logNum, y = avgGSW, color = cycle)) +
  geom_line(data = LITUfresh, aes(x = logNum, y = avgGSW, color = cycle)) + 
  geom_line(data = LITUsalt, aes(x = logNum, y = avgGSW, color = cycle)) +
  geom_point(data = LITUcontrolA, aes(x = logNum, y = avgGSW, color = cycle)) +
  geom_point(data = LITUcontrolB, aes(x = logNum, y = avgGSW, color = cycle)) + 
  geom_point(data = LITUbucket, aes(x = logNum, y = avgGSW, color = cycle)) +
  geom_point(data = LITUdetached, aes(x = logNum, y = avgGSW, color = cycle)) +
  geom_point(data = LITUfresh, aes(x = logNum, y = avgGSW, color = cycle)) + 
  geom_point(data = LITUsalt, aes(x = logNum, y = avgGSW, color = cycle)) +
  theme_bw() +
  ggtitle("all together litu") 


FAGR_A_avgA <- plotAvgA(FAGRcontrolA, FAGRbucket, FAGRdetached)
FAGR_A_avgA
FAGR_B_avgA <- plotAvgA(FAGRcontrolB, FAGRfresh, FAGRsalt)
FAGR_B_avgA

FAGR_A_avgGSW <- plotAvgGSW(FAGRcontrolA, FAGRbucket, FAGRdetached, upper = 0.1)
FAGR_A_avgGSW
FAGR_B_avgGSW <- plotAvgGSW(FAGRcontrolB, FAGRfresh, FAGRsalt, upper = 0.1)
FAGR_B_avgGSW

ggplot() + 
  geom_line(data = FAGRcontrolA, aes(x = logNum, y = avgA, color = cycle)) +
  geom_line(data = FAGRcontrolB, aes(x = logNum, y = avgA, color = cycle)) + 
  geom_line(data = FAGRbucket, aes(x = logNum, y = avgA, color = cycle)) +
  geom_line(data = FAGRdetached, aes(x = logNum, y = avgA, color = cycle)) +
  geom_line(data = FAGRfresh, aes(x = logNum, y = avgA, color = cycle)) + 
  geom_line(data = FAGRsalt, aes(x = logNum, y = avgA, color = cycle)) +
  geom_point(data = FAGRcontrolA, aes(x = logNum, y = avgA, color = cycle)) +
  geom_point(data = FAGRcontrolB, aes(x = logNum, y = avgA, color = cycle)) + 
  geom_point(data = FAGRbucket, aes(x = logNum, y = avgA, color = cycle)) +
  geom_point(data = FAGRdetached, aes(x = logNum, y = avgA, color = cycle)) +
  geom_point(data = FAGRfresh, aes(x = logNum, y = avgA, color = cycle)) + 
  geom_point(data = FAGRsalt, aes(x = logNum, y = avgA, color = cycle)) +
  theme_bw() +
  ggtitle("all together fagr") 

ggplot() + 
  geom_line(data = FAGRcontrolA, aes(x = logNum, y = avgGSW, color = cycle)) +
  geom_line(data = FAGRcontrolB, aes(x = logNum, y = avgGSW, color = cycle)) + 
  geom_line(data = FAGRbucket, aes(x = logNum, y = avgGSW, color = cycle)) +
  geom_line(data = FAGRdetached, aes(x = logNum, y = avgGSW, color = cycle)) +
  geom_line(data = FAGRfresh, aes(x = logNum, y = avgGSW, color = cycle)) + 
  geom_line(data = FAGRsalt, aes(x = logNum, y = avgGSW, color = cycle)) +
  geom_point(data = FAGRcontrolA, aes(x = logNum, y = avgGSW, color = cycle)) +
  geom_point(data = FAGRcontrolB, aes(x = logNum, y = avgGSW, color = cycle)) + 
  geom_point(data = FAGRbucket, aes(x = logNum, y = avgGSW, color = cycle)) +
  geom_point(data = FAGRdetached, aes(x = logNum, y = avgGSW, color = cycle)) +
  geom_point(data = FAGRfresh, aes(x = logNum, y = avgGSW, color = cycle)) + 
  geom_point(data = FAGRsalt, aes(x = logNum, y = avgGSW, color = cycle)) +
  theme_bw() +
  ggtitle("all together fagr") 


```







